### 2017-01-12 domain_level/domain_tree_analysis.Rmd
initial commit

---
title: "Domain Tree Analysis"
author: "Gerry Tonkin-Hill"
date: "11/11/2015"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

##Load libraries
```{r}
library(ggplot2)
library(reshape2)
library(edgeR)
library(plyr)
library(scales)

library(Rsubread)
library(structSSI)
library(data.table)
library(igraph)

library(cluster)
library(fastcluster)
library(RFLPtools)

library(pheatmap)
library(EDASeq)
library(dynamicTreeCut)
library(moduleColor)
library(knitr)

library(DESeq2)

wd = "./"
colors <- c("#0571b0","#ca0020")
```

##Get read counts for each domain
Used featureCounts allowing for overlap. That is, if a read overlaps more than one segment it is counted as a hit for both segments.
```{r}
setwd(wd)
# files <- Sys.glob("/home/users/allstaff/tonkin-hill.g/gene_level_var_analysis/alignments_seperate_assemblies_best1/*.sam")
# dom_counts <- featureCounts(files, annot.ext = "./data/NTS_DBL_CIDR_seperate_assembly_domains.SAF"
#                             , isPairedEnd=TRUE, isGTFAnnotationFile=FALSE
#                             , countChimericFragments=TRUE
#                             , allowMultiOverlap=TRUE
#                             , useMetaFeatures=FALSE
#                             , nthreads=20
#                             , countMultiMappingReads=TRUE)
# 
# #combine counts for different runs of the same sample
# colnames(dom_counts$counts) <- gsub(".*_best1\\.","",colnames(dom_counts$counts))
# colnames(dom_counts$counts) <- gsub("_S.*","",colnames(dom_counts$counts))
# temp <- dom_counts
# temp$counts <- t(rowsum(t(temp$counts), group = rownames(t(temp$counts))))
# dim(temp$counts)
# dom_counts <- temp
# # #
# saveRDS(dom_counts, "./data/dom_counts.rds")

dom_counts <- readRDS("./data/dom_counts.rds")
```

Again we remove samples SFD1, SFM9 and SFC.025 as they were not sequenced to a sufficient depth.
```{r}
setwd(wd)
dom_counts$counts <- dom_counts$counts[,!(colnames(dom_counts$counts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
```

##Investigate previous domain classifications
```{r}
setwd(wd)
prev_dom_counts <- data.frame(dom_counts$counts)
prev_dom_counts$domainType <- unlist(lapply(rownames(prev_dom_counts), function(x){
  temp <- unlist(strsplit(x, "_"))
  return(temp[length(temp)-1])
  }))
# prev_dom_counts$domainType <- gsub("\\..*", "", prev_dom_counts$domainType)
prev_dom_counts <- ddply(prev_dom_counts, "domainType", numcolwise(sum))
rownames(prev_dom_counts) <- prev_dom_counts$domainType
prev_dom_counts$domainType <- NULL
prev_dom_counts <- as.matrix(prev_dom_counts)

x_domains <- DGEList(counts=prev_dom_counts)
x_domains <- calcNormFactors(x_domains, method="RLE")
categories <- as.factor(substring(rownames(x_domains$samples),1,1))
```
A total of `r dim(x_domains)[1]` domains from the those of Rask et al are identified in the samples.

##A PCA plot of normalised counts for domains of Rask et al
Reads are annotated to the de novo assembled transcripts which are annotated with the hidden Markov model domains profiles of Rask et al.
```{r}
plotPCA(cpm(x_domains,normalized.lib.sizes=TRUE, log=T, prior.count=0.5), col=colors[categories], cex=0.8, isLog=TRUE)
title("VAR PCA - Previous Domains")
```

##Investigate differential expression
```{r}
colData <- data.frame(disease=categories)
rownames(colData) <- colnames(x_domains$counts)
dds_dom <- DESeqDataSetFromMatrix(countData = x_domains$counts
                              , colData = colData
                              , design = ~disease)
dds_dom <- estimateSizeFactors(dds_dom)
dds_dom <- DESeq(dds_dom) 
res <- results(dds_dom,contrast=c("disease","S","I"), pAdjustMethod = "BH")
summary(res, alpha=0.05)
```

##A heatmap of the Up-regulated domains
```{r}
res_domains <- res
resOrdered <- res_domains[order(res_domains$padj),]
resSig <- subset(resOrdered, padj < 0.05)

resSigLFC <- data.frame(subset(resSig, log2FoldChange > 0))

select <- order(rowMeans(counts(dds_dom, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds_dom)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resSigLFC),]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe" 
annotData$phenotype[annotData$disease=="S"] <- "severe"
# annotData <- annotData[annotData$disease=="S",]
annotData$disease <- NULL
pheatmap(log2.norm.counts, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=TRUE, annotation_col=annotData, annotation_colors = anno_colors)
```

##Boxplot of NTSA and NTSB domains
```{r}
setwd(wd)

select <- order(rowMeans(counts(dds_dom, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds_dom)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% c("NTSA", "NTSB"),]

normCounts <- data.frame(t(log2.norm.counts))
normCounts$Sample <- rownames(normCounts)
normCounts$phenotype[as.factor(substring(rownames(normCounts),1,1))=='S'] <-  "severe"
normCounts$phenotype[as.factor(substring(rownames(normCounts),1,1))=='I'] <-  "non-severe"
normCounts <- melt(normCounts, id.vars=c("Sample","phenotype"), variable.name="Type", value.name="Normalised.Count")

gg <- ggplot(normCounts, aes(x=Type, y=Normalised.Count, colour=phenotype))
# gg <- gg + geom_point(position=position_jitter(w=0.1,h=0))
gg <- gg + geom_boxplot()
gg <- gg + theme_bw()
gg <- gg + scale_colour_manual(values = colors)
gg
```

##Get clusters
Hierarchical clustering is performed by iteratively running **usearch** at identity levels 0.97,0.95,0.9,0.85,0.8,0.75,0.7,0.65,0.6,0.5 with the command

*usearch -cluster_fast input.fasta -sort length -wordlength 2 -fulldp -maxaccepts 0 -maxrejects 0 -alpha aa -mincols -id i -centroids out.fasta -uc clusters.uc*

This was chosen as it effectively dealt with fragments of domains. The method is **not** guaranteed to produce clusters where every sequence in the cluster is within the identity threshold of the centroid.

```{r}
setwd(wd)
clusters <- read.csv("./data/NTS_DBL_CIDR_seperate_assembly_domains_PreProcessed_hierarchical_clusters.csv")

stopifnot(nrow(dom_counts$counts)==nrow(clusters))

dom_counts_clusters <- merge(dom_counts$counts, clusters, by.x=0 , by.y="Transcript")
```

##Create hierarchical structure

Starting with no clustering, individual domain counts are combined at the different hierarchical clustering levels.

```{r}
setwd(wd)

#first create an edgelist
edges <- list()
edges <- rbind(edges,cbind("root", paste(as.character(clusters[,2]), colnames(clusters)[2], sep="_")))
for (j in 3:(ncol(clusters))){
      edges <- rbind(edges,cbind(paste(as.character(clusters[,j-1]), colnames(clusters)[j-1], sep="_"), paste(as.character(clusters[,j]), colnames(clusters)[j], sep="_")))
  }
edges <- edges[!duplicated(edges),]
edges <- apply(as.matrix(edges), 2,as.character)

#now sum counts at the different clustering heights
results <- data.frame()
pca.df <- data.frame()
all_clusters.matrix <- data.frame()
for (c in colnames(clusters)[2:(ncol(clusters))]) {
  #reduce matrix to cluster at level c
  dom_counts_clusters.dt <- data.table(dom_counts_clusters)
  dom_counts_clusters.dt <- dom_counts_clusters.dt[,lapply(.SD, sum, na.rm=TRUE)
                                                       , by=c
                                                       , .SDcols=colnames(dom_counts_clusters)[2:42]]

  dom_counts_clusters.matrix <- as.matrix(dom_counts_clusters.dt)
  dom_counts_clusters.matrix <- dom_counts_clusters.matrix[,2:ncol(dom_counts_clusters.matrix)]
  rownames(dom_counts_clusters.matrix) <- paste(dom_counts_clusters.dt[[1]], c, sep="_")
  all_clusters.matrix <- rbind(all_clusters.matrix, dom_counts_clusters.matrix)

  #generate points for PCA
  dgePCA <- DGEList(counts=dom_counts_clusters.matrix)
  colData <- data.frame(disease=as.factor(substring(colnames(dom_counts_clusters.matrix),1,1)))
  rownames(colData) <- rownames(dgePCA$samples)
  ddsPCA <- DESeqDataSetFromMatrix(countData = dgePCA$counts#dom_counts_clusters.matrix
                            , colData = colData
                            , design = ~disease)
  cts <- counts(ddsPCA)
  geoMeans <- apply(cts, 1, function(row) if (all(row == 0)) 0 else exp(mean(log(row[row != 0]))))
  ddsPCA <- estimateSizeFactors(ddsPCA, geoMeans=geoMeans)
  pca <- plotPCA(varianceStabilizingTransformation(ddsPCA), intgroup=c("disease"), returnData = TRUE)
  pca$level <- c
  pca.df <- rbind(pca.df, pca)
}
```

##PCA plots at different levels
```{r}
setwd(wd)
pca.df$phenotype[pca.df$disease=='S'] <- "severe"
pca.df$phenotype[pca.df$disease=='I'] <- "non-severe"
gg <- ggplot(pca.df, aes(PC1, PC2, color=phenotype)) + scale_color_manual(values=colors)
# gg <- gg + geom_text(aes(label=name))
gg <- gg + geom_point()
gg <- gg + facet_wrap(~level)
gg <- gg + theme_bw() 
gg
```

##A PCA plot of normalised counts for domains clustered at 50% identity
```{r}
setwd(wd)

dom_counts_clusters.dt <- data.table(dom_counts_clusters)
dom_counts_clusters.dt <- dom_counts_clusters.dt[,lapply(.SD, sum, na.rm=TRUE)
                                                       , by='X0.5'
                                                       , .SDcols=colnames(dom_counts_clusters)[2:42]]

dom_counts_clusters.matrix <- as.matrix(dom_counts_clusters.dt)
dom_counts_clusters.matrix <- dom_counts_clusters.matrix[,2:ncol(dom_counts_clusters.matrix)]
rownames(dom_counts_clusters.matrix) <- paste(dom_counts_clusters.dt[[1]], c, sep="_")

#generate points for PCA
my_domains <- DGEList(counts=dom_counts_clusters.matrix)
my_domains <- calcNormFactors(my_domains, method="RLE")
plotPCA(cpm(my_domains,normalized.lib.sizes=TRUE, log=T, prior.count=0.5), col=colors[categories], cex=0.8, isLog=TRUE)
title("VAR PCA - Domains Clustered at 50% idenitiy")
```

##Differential expression testing

First, to speed up computation and focus on only the most interesting domain clusters we remove those with less than 15 reads per million which appear in less than 5 samples.

```{r}
setwd(wd)

dge_clust <- DGEList(counts=as.matrix(all_clusters.matrix))

#create design matrix
colData <- data.frame(disease=as.factor(substring(rownames(dge_clust$samples),1,1)))
rownames(colData) <- rownames(dge_clust$samples)

cpm(10, mean(dge_clust$samples$lib.size))

keep <- rowSums(cpm(dge_clust)>15) >= 5
dge_clust <- dge_clust[keep,,keep.lib.sizes=FALSE]
```

DESeq2 is now run with default parameters using the Benjamini-Yeukatelli procedure for multiple testing correction. We only take one cluster from each path through the tree (root to leaf) and thus overall we are more conservative than conventional Benjamini-Yeukatelli.

```{r, cache=FALSE}
setwd(wd)

#create DESeq2 dataset
dds <- DESeqDataSetFromMatrix(countData = dge_clust$counts
                            , colData = colData
                            , design = ~disease)

#estimate size factors
dds <- estimateSizeFactors(dds)

#run DESeq2 pipeline
dds <- DESeq(dds)
results <- results(dds,contrast=c("disease","S","I"), pAdjustMethod = "BY")
summary(results, alpha=0.05)
```

Match up p-values with edges in the tree structure and plot trees that have a significant node.

```{r}
setwd(wd)

Sig <- results[!is.na(results$padj),]
Sig <- Sig[Sig$padj<0.05,]

tree <- graph.edgelist(edges)

naNodes <- V(tree)$name[!(V(tree)$name %in% rownames(results))]
na1s = rep(1, length(naNodes))

unadjp <- c(na1s, results$pvalue)
adjp <- c(rep(NA, length(naNodes)), results$padj)
names(unadjp) <- c(naNodes, rownames(results))
names(adjp) <- c(naNodes, rownames(results))

if (!all(names(unadjp) %in% unique(as.vector(edges)))) {
        stop("Names of elements in unadjp do not match names of tree nodes")
    }
if (!all(unique(as.vector(edges)) %in% names(unadjp))) {
        stop("Names of elements in unadjp do not match names of tree nodes")
    }

p.vals <- data.frame(unadjp, adjp = adjp)
rownames(p.vals) <- names(unadjp)

cluster_names <- clusters
for (i in 2:ncol(clusters)){
  cluster_names[,i] <- paste(as.character(cluster_names[,i]), colnames(clusters)[i], sep="_")
}
```

##Functions to explore the tree

These function return the ancestors and children of a given node in the hierarchical tree.

```{r}
setwd(wd)

#return the children of a node
get_children <- function(node, edges) {
    if (is.null(node)){
      return(NULL)
    }
    tree.el.tmp <- data.frame(parent = edges[, 1], child = edges[, 2], stringsAsFactors = F)
    children <- tree.el.tmp[which(tree.el.tmp$parent == node), "child"]
    children <- unlist(c(children, lapply(children, get_children, edges)))
    return(children)
}

#return the anscestors of a node
get_parents <- function(node, edges){
  if (is.null(node)){
      return(NULL)
    }
    tree.el.tmp <- data.frame(parent = edges[, 1], child = edges[, 2], stringsAsFactors = F)
    parent <- tree.el.tmp[which(tree.el.tmp$child == node), "parent"]
    parent <- unlist(c(parent, lapply(parent, get_parents, edges)))
    return(parent)
}

#plot the subtree of a node
plot_tree <- function(node, edges, p.vals, return_script=FALSE){
  tree <- graph.edgelist(edges)
  ancestors <- get_parents(node, edges)
  root <- ancestors[length(ancestors)-1] #choose root one level below root of entire tree
  if (ancestors[1]=="root"){
    root <- node
  }

  keep <- c(get_children(root, edges), root)
  newTree <- induced.subgraph(tree, vids=V(tree)[V(tree)$name %in% keep])
  p.vals <- p.vals[rownames(p.vals) %in% V(newTree)$name,]
  p.vals <- p.vals[match(unique(as.vector(t(get.edgelist(newTree)))), rownames(p.vals)),]

  hyp.tree <- new("hypothesesTree", tree = get.edgelist(newTree),
        p.vals = p.vals, alpha = 0.05)

  plot(hyp.tree, adjust = TRUE, return_script=return_script)
}
```

##Get significant nodes in trees

From those nodes that are found to be significant after multiple testing correction we keep only the most signifcant from one path through the tree. That is if a node has an ancestor or child that has a smaller p-value, it will not be considered. This is achieved by iteratively selecting the most significant node and removing it's ancestors and children from the list of signifcant nodes.

```{r}
setwd(wd)

tempSig <- Sig
tempSig$cluster <- as.numeric(gsub(".*_X","",rownames(tempSig)))
tempSig <- tempSig[order(tempSig$cluster),]
SigFilteredResults <- data.frame()
while (nrow(tempSig)>0){
  trow <- tempSig[which.min(tempSig$padj),]
  SigFilteredResults <- rbind(SigFilteredResults, trow)
  t <- rownames(trow)
  tChildren <- get_children(t, edges)
  tParents <- get_parents(t, edges)
  tempSig <- tempSig[!(rownames(tempSig) %in% c(t, tChildren, tParents)),]
  tempSig <- tempSig[order(tempSig$cluster),]
}

resUp <- SigFilteredResults[SigFilteredResults$log2FoldChange>0,]
resDown <- SigFilteredResults[SigFilteredResults$log2FoldChange<0,]
```

Overall we have found `r nrow(SigFilteredResults)` significantly DE clusters of which `r sum(SigFilteredResults$log2FoldChange>0)` are up-regulated.

Update final results to include information on the transcripts that contain them
```{r}
#Collect sequence labels for each cluster
final <- data.frame(SigFilteredResults)
sig_clusters <- as.numeric(gsub("_.*","", rownames(SigFilteredResults)))
levels <- gsub(".*_","", rownames(SigFilteredResults))
significantClusters <- list()
for (i in 1:length(levels)){
  significantClusters[[i]] <- clusters[clusters[,colnames(clusters)==levels[i]]==sig_clusters[i],]$Transcript
}
final$sequences <- I(significantClusters)

getDomain <- function(x){
  x <- x[[1]][1]
  x <- toString(x)
  x <- strsplit(x, "_")
  x <- x[[1]]
  x <- x[length(x)-1]
  return(x)
}
cluster_domain_type <- unlist(lapply(final$sequences, getDomain))
names(cluster_domain_type) <- rownames(final)
```

##Clustering significant domains based on co-expression
```{r}
# expression <- cpm(dge_clust, normalized.lib.sizes=TRUE, log=TRUE)[rownames(dge_clust) %in% rownames(resUp),]
select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
expression <- assay(nt)[rownames(nt) %in% rownames(resUp),]

dissim = dist(expression)
dendro = hclust(dissim, method = "complete")

AutoColor = NULL
for (deepSplit in 0:3)
  AutoColor = cbind(AutoColor, labels2colors(cutreeDynamic(dendro = dendro,
                    minClusterSize = 2,
                    cutHeight = NULL, method = "hybrid", deepSplit = deepSplit,
                    distM = as.matrix(dissim))))
AutoLabels = paste("Hybrid 'auto': dS =", c(0:3))

cut <- cutreeDynamic(dendro = dendro,
                    minClusterSize = 2,
                    cutHeight = NULL, method = "hybrid", deepSplit = 0,
                    distM = as.matrix(dissim))

rowLabels <- dendro$labels
# rowLabels <- rowLabels[dendro$order]
rowsAnnot <- data.frame(group=labels2colors(cut))
rownames(rowsAnnot) <- rowLabels

par(mfrow=c(2,1))
par(cex = 1.4);
par(mar = c(0,8.5,2,0));
plot(dendro,labels=F,main="Co-expression clustering with dynamic tree cut", sub="", xlab="")
par(mar = c(1,8.5,0,0));
plotHclustColors(dendro, AutoColor, 
            AutoLabels, 
            main = "")
```

##Heatmap of up-regulated clusters

```{r}
setwd(wd)

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe"
annotData$phenotype[annotData$disease=="S"] <- "severe"
annotData$disease <- NULL

log2.norm.counts <- log2.norm.counts[dendro$order,]
pheatmap(log2.norm.counts, cluster_rows=FALSE, show_rownames=TRUE, cluster_cols=TRUE
         , annotation_col=annotData, annotation_colors = anno_colors
         , clustering_method = "complete", cutree_rows=9
         , labels_row=paste(rownames(log2.norm.counts), cluster_domain_type[rownames(log2.norm.counts)] ,sep=" - ")
         , annotation_row=rowsAnnot, fontsize_row=7)
```

##Heatmap of down-regulated clusters

```{r}
setwd(wd)

nt <- normTransform(dds)
expression <- assay(nt)[rownames(nt) %in% rownames(resDown),]
dissim = dist(expression)
dendro = hclust(dissim, method = "complete")
cut <- cutreeDynamic(dendro = dendro,
                    minClusterSize = 2,
                    cutHeight = NULL, method = "hybrid", deepSplit = 0,
                    distM = as.matrix(dissim))
rowLabels <- dendro$labels
rowsAnnot <- data.frame(group=labels2colors(cut))
rownames(rowsAnnot) <- rowLabels

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resDown),]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe"
annotData$phenotype[annotData$disease=="S"] <- "severe"
annotData$disease <- NULL

log2.norm.counts <- log2.norm.counts[dendro$order,]
pheatmap(log2.norm.counts, cluster_rows=FALSE, show_rownames=TRUE, cluster_cols=TRUE
         , annotation_col=annotData, annotation_colors = anno_colors
         , clustering_method = "complete", cutree_rows=10
         , labels_row=paste(rownames(log2.norm.counts), cluster_domain_type[rownames(log2.norm.counts)] ,sep=" - ")
         , annotation_row=rowsAnnot, fontsize_row=6)
```

##PCA plot plot of significant clusters

```{r}
setwd(wd)
plotPCA(cpm(dge_clust, normalized.lib.sizes=TRUE, log=T)[rownames(dge_clust) %in% rownames(SigFilteredResults),], col=colors[colData$disease], cex=0.8, isLog=TRUE)
```

##Table of results
```{r}
setwd(wd)

#Now print the table
final$padj <- format(final$padj, scientific=TRUE, digits=3)
final$pvalue <- format(final$pvalue, scientific=TRUE, digits=3)
final$sequences <- unlist(lapply(final$sequences, paste, collapse=","))
kable(final, digits=3)
```

##Create Tree Plots
```{r}
# writeTree <- function(node, directory, edges, p.vals){
#     sink(paste(paste(directory,node,sep=""), "_tree.html", sep=""))
#     cat(plot_tree(node, edges, p.vals, return_script=TRUE))
#     sink()
#     closeAllConnections()
# #     close(paste(paste(directory,node,sep="_"), "_tree.html"))
# }
# plot_tree("670_X0.6", edges, p.vals, return_script=F)
# writeTree("670_X0.6", "/home/users/allstaff/tonkin-hill.g/Dropbox/Var transcriptome manuscript/Figures/VAR differential analysis (4)/domainTrees/", edges, p.vals)
# lapply(rownames(final), writeTree, "/home/users/allstaff/tonkin-hill.g/var_differential_analysis/domain_level/tree_figures/", edges, p.vals)
```

##Create Dot Plots for Tree Diagrams
```{r}
# plotExpression <- function(cluster, expression, results, outdir){
#   clustExp <- data.frame(expression[rownames(expression)==cluster,])
#   clustExp$Phenotype[as.factor(substring(rownames(clustExp),1,1))=='S'] <-  "severe"
#   clustExp$Phenotype[as.factor(substring(rownames(clustExp),1,1))=='I'] <-  "non-severe"
#   colnames(clustExp) <- c("expression", "phenotype")
#   clustRes <- results[rownames(results)==cluster,]
#   gg <- ggplot(clustExp, aes(x = factor(phenotype), y = expression, fill=factor(phenotype)))
#   gg <- gg + geom_dotplot(binaxis = "y", stackdir = "center", binpositions="all")
#   gg <- gg + theme_bw() + theme(legend.position="none")
#   gg <- gg + scale_fill_manual(values = colors)
#   gg <- gg + labs(title = paste(c("Cluster:", cluster
#                              , "P-value:", format(clustRes$padj, scientific=TRUE, digits=3)
#                              , "LFC:", format(clustRes$log2FoldChange, scientific=TRUE, digits=3))
#                              , sep="", collapse="  ")
#                     , x = "Scaled expression"
#                     , y = "Phenotype"
#                     )
# #   print(gg)
#   ggsave(filename=paste(c(outdir, cluster, "_expression.png"),sep="",collapse=""), plot=gg)
#   }
# 
# 
# 
# lapply(c("91_X0.5", get_parents("91_X0.5", edges), get_children("91_X0.5", edges)), plotExpression, expression, results, "/home/users/allstaff/tonkin-hill.g/var_differential_analysis/domain_level/tree_figures/expression_dot_plots/")

# write.table(results, file="/home/users/allstaff/tonkin-hill.g/var_differential_analysis/domain_level/tree_figures/results.csv", sep=",", quote=F)
# clust <- data.frame(clusters)
# clust$transcript <- clust$Transcript
# clust$Transcript <- NULL
# write.table(clust, file="/home/users/allstaff/tonkin-hill.g/var_differential_analysis/domain_level/tree_figures/segement_clusters_complete.csv", sep=",", quote=F, row.names=F)
#                   
# get_root <- function(node, edges){
#   parents = get_parents(node, edges)
#   if (length(parents)==1){
#     return(node)mai
#   }
#   else{
#     return(parents[length(parents)-1])
#   }
# }
# 
# resSig <- results[!is.na(results$padj),]
# resSig <- resSig[resSig$padj<=0.05,]
# 
# sig <- data.frame(clusters=rownames(SigFilteredResults), root=gsub("_.*", "", unlist(lapply(rownames(SigFilteredResults), get_root, edges))))
# write.table(sig, file="/home/users/allstaff/tonkin-hill.g/var_differential_analysis/domain_level/tree_figures/sig_clusters.csv", sep=",", row.names=FALSE, quote = FALSE)

```

### 2017-01-12 segment_level/scripts/combine_gismo_clusters_generate_saf.py

@@ -1,0 +1,28 @@
+import glob
+
+mult=1
+with open("combined_blocks_clustered.csv", 'w') as outfile:
+  for f in glob.glob("*hierarchical_clusters.csv"):
+    print f
+    with open(f, 'rU') as infile:
+      if mult==1:
+        outfile.write(infile.next())
+      else:
+        infile.next()
+      for line in infile:
+        line = line.strip().split(",")
+        outfile.write(",".join([line[0]] + [str(int(t)+10000*mult) for t in line[1:]])+"\n")
+    mult+=1
+
+mult=0
+with open("combined_blocks.SAF", 'w') as outfile:
+  for f in glob.glob("*safile.SAF"):
+    mult+=1
+    print f
+    with open(f, 'rU') as infile:
+      if mult==1:
+        outfile.write(infile.next())
+      else:
+        infile.next()
+      for line in infile:
+        outfile.write(line)

### 2017-01-12 segment_level/scripts/split_gismo_alignment_into_blocks.py

@@ -1,0 +1,372 @@
+import sys,os
+import argparse
+from subprocess import check_call
+# from mungo.fasta import FastaReader
+import re
+import numpy as np
+from collections import defaultdict
+from itertools import groupby
+from mungo.sequence import sixFrameTranslation
+
+ACCEPTABLE_SYMBOLS = ['*', '-', 'A', 'C', 'D', 'E', 'F', 'G', 'H', 'I', 'K', 'L', 'M', 'N', 'P', 'Q', 'R', 'S', 'T', 'V', 'W', 'Y']
+ACCEPTABLE_SYMBOLS_SET = set(ACCEPTABLE_SYMBOLS)
+
+class writeable_dir(argparse.Action):
+  def __call__(self,parser, namespace, values, option_string=None):
+    prospective_dir=values
+    if not os.path.isdir(prospective_dir):
+        raise argparse.ArgumentTypeError("writeable_dir:{0} is not a valid path".format(prospective_dir))
+    if os.access(prospective_dir, os.W_OK):
+        setattr(namespace,self.dest,prospective_dir)
+    else:
+        raise argparse.ArgumentTypeError("writeable_dir:{0} is not a writeable dir".format(prospective_dir))
+
+class readable_file(argparse.Action):
+  def __call__(self,parser, namespace, values, option_string=None):
+    for v in values:
+      prospective_file=v
+      if not os.path.isfile(prospective_file):
+          raise argparse.ArgumentTypeError("readable_dir:{0} is not a valid file path".format(prospective_file))
+      if os.access(prospective_file, os.R_OK):
+          setattr(namespace,self.dest,prospective_file)
+      else:
+          raise argparse.ArgumentTypeError("readable_dir:{0} is not a readable file".format(prospective_file))
+
+class MyError(Exception):
+    def __init__(self, value):
+        self.value = value
+    def __str__(self):
+        return repr(self.value)
+
+
+def readFasta(fastafile):
+
+  with open(fastafile, 'rU') as fp:
+    name, seq = None, []
+    for line in fp:
+        line = line.rstrip()
+        if line.startswith(">"):
+            line=line[1:]
+            if name: yield (name, ''.join(seq))
+            name, seq = line, []
+        else:
+            seq.append(line)
+    if name: yield (name, ''.join(seq))
+
+def checkSeq(test_str):
[…]

### 2017-01-12 domain_level/tree_comparison_with_Bengtsson_and_Rask.Rmd
Added in domain comparisons with bengtsson and rask

---
title: "Supplementary domain trees"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

##Load libraries
```{r}
library(ggtree)
library(ggplot2)
library(ape)
wd = "/home/users/allstaff/tonkin-hill.g/var_differential_analysis/domain_level/data"
```

#Phylogentic method
Protein domains were aligned using MUSCLE v3.8.31 and phylogentic trees were constructed using FastTree version 2.1.7 SSE3
```
muscle -in domains.fasta -out MA_domains.fasta
FastTree < MA_domains.fasta > MA_domains.tree
```

#DC5 Comparison
Bengtsson et al. found that DC4 may be related to PfEMP1 proteins binding to ICAM-1. In particular they identified a ditinct clade of DBLb3 domains that bound to ICAM-1. The tree drawn below indicates that the first DBLb3 domain of PFD125w (one of those found in the clade) clusters closest to the severe cluster DBLb.s.2

Lavsten et al 2012 also found DBLb3 to be more expressed in children with severe malaria.
```{r}
setwd(wd)

dbl_b3 <- read.tree("./compare_DBLb3/combined_vardom_severe_DBLb3.phylip")

labels <- "black"
labels[grepl("\\.s\\.1", dbl_b3$tip.label)] <- "#4daf4a"
labels[grepl("\\.s\\.2", dbl_b3$tip.label)] <- "#f781bf"
labels[grepl("PFD1235w", dbl_b3$tip.label)] <- "#e41a1c"

plot.phylo(dbl_b3, type="phylogram",
            use.edge.length = FALSE,
           show.node.label = TRUE,
           cex=0.5,
           tip.color=labels,
           align.tip.label=TRUE)
legend(1, -1, legend=c("DBLb3", "DBLb3.s.1", "DBLb3.s.2", "PFD1235w - Bengsston"),
                   fill=c("black", "#4daf4a" ,"#f781bf","#e41a1c"), 
       xpd=NA,
       cex=0.5)
```

#DC8 Comparison

Domain cassette 8, first decribed by Rask et al has been associated with severe forms of malaria. Here, the severe clusters identified in the domain analysis are clustered with those from domain cassette 8 of Rask et al. Specifically we look at DBLb12.s.1, DBLd1.s.1-4 and CIDRb1.s.1

##DBLd1
The DBLd1 domains from DC8 containing proteins of Rask et al are failry well distributed amoungst the remaining DBLd1 domains of Rask et al. Notably, two DC8 DBLd1 domains cluster closely with severe domain clusters DBLd1.s.3 and DBLd1.s.4.

Severe clusters DBLd1.s.1 and DBLd1.s.2 also have DC8 DBLd1 domains nearby but are more closely clustered to a number of non DC8 DBLd1 domains.

```{r}
setwd(wd)

tree <- read.tree("./compare_DC8/MA_combined_raskDBLd1_DBLd1s.fasta.phy")

labels <- "black"
labels[grepl("\\.s\\.1", tree$tip.label)] <- "#f781bf"
labels[grepl("\\.s\\.2", tree$tip.label)] <- "#984ea3"
labels[grepl("\\.s\\.3", tree$tip.label)] <- "#cab2d6"
labels[grepl("\\.s\\.4", tree$tip.label)] <- "#ff7f00"
labels[grepl("^DC8", tree$tip.label)] <- "#e41a1c"

plot.phylo(tree, type="phylogram",
            use.edge.length = FALSE,
           show.node.label = TRUE,
           cex=0.5,
           tip.color=labels,
           align.tip.label=TRUE)
legend(1, -1, legend=c("DBLd1", "DBLd1.s.1", "DBLd1.s.2", "DBLd1.s.3", "DBLd1.s.4", "DBLd1 from DC8"),
                   fill=c("black", "#f781bf","#984ea3","#cab2d6","#ff7f00","#e41a1c"), 
       xpd=NA,
       cex=0.5)
```

##DBLb12
The tree below is not very informative due to the very high level of conservation found in the de novo assembled DBLb12 transcripts. However the DBLb12 transcripts from PFEMP1 domains of Rask et al split into two groups which are equally distant from each other as they are from the tightly clustered severe group. This suggests that the severe DBLb12 cluster could be grouped with those from DC8 as there is just as much variation between DC8 DBLb12 domains as there is between the severe group and DC8 DBLb12 domains.

```{r}
setwd(wd)

tree <- read.tree("./compare_DC8/MA_combined_raskDBLb12_DBLb12s.fasta.phy")

labels <- "black"
labels[grepl("\\.s\\.1", tree$tip.label)] <- "#ff7f00"
labels[grepl("^DC8", tree$tip.label)] <- "#e41a1c"

plot.phylo(tree, type="phylogram",
            use.edge.length = FALSE,
           show.node.label = TRUE,
           cex=0.5,
           tip.color=labels,
           align.tip.label=TRUE)
legend(1, -1, legend=c("DBLb12", "DBLb12", "DBLb12 from DC8"),
                   fill=c("black", "#ff7f00","#e41a1c"), 
       xpd=NA,
       cex=0.5)
```

##CIDRb1

The severe CIDRb1 domain clusters together with one DC8 CIDRb1 domain from Rask et al. The remaining CIDRb1 domains from proteins that conatin DC8 in Rask et al are distributed across the remianing CIDRb1 domains from Rask et al and don't form a disctinct group.

```{r}
setwd(wd)

tree <- read.tree("./compare_DC8/MA_combined_raskCIDRb1_CIDRb1s.fasta.phy")

labels <- "black"
labels[grepl("\\.s\\.1", tree$tip.label)] <- "#ff7f00"
labels[grepl("^DC8", tree$tip.label)] <- "#e41a1c"

plot.phylo(tree, type="phylogram",
            use.edge.length = FALSE,
           show.node.label = TRUE,
           cex=0.5,
           tip.color=labels,
           align.tip.label=TRUE)
legend(1, -1, legend=c("CIDRb1", "CIDRb1.s.1", "CIDRb1 from DC8"),
                   fill=c("black", "#ff7f00","#e41a1c"), 
       xpd=NA,
       cex=0.5)
```

###System Info
```{r}
sessionInfo()
```

### 2017-02-17 domain_level/tree_comparison_with_Bengtsson_and_Rask.Rmd
added in html updated domain rmd to include session info

@@
 ```{r}
-library(ggtree)
 library(ggplot2)
 library(ape)
-wd = "/home/users/allstaff/tonkin-hill.g/var_differential_analysis/domain_level/data"
 ```
@@
 ```{r}
-setwd(wd)
-
-dbl_b3 <- read.tree("./compare_DBLb3/combined_vardom_severe_DBLb3.phylip")
+dbl_b3 <- read.tree("./data/compare_DBLb3/combined_vardom_severe_DBLb3.phylip")
 
@@
 ```{r}
-setwd(wd)
-
-tree <- read.tree("./compare_DC8/MA_combined_raskDBLd1_DBLd1s.fasta.phy")
+tree <- read.tree("./data/compare_DC8/MA_combined_raskDBLd1_DBLd1s.fasta.phy")
 
@@
 ```{r}
-setwd(wd)
-
-tree <- read.tree("./compare_DC8/MA_combined_raskDBLb12_DBLb12s.fasta.phy")
+tree <- read.tree("./data/compare_DC8/MA_combined_raskDBLb12_DBLb12s.fasta.phy")
 
@@
 ```{r}
-setwd(wd)
-
-tree <- read.tree("./compare_DC8/MA_combined_raskCIDRb1_CIDRb1s.fasta.phy")
+tree <- read.tree("./data/compare_DC8/MA_combined_raskCIDRb1_CIDRb1s.fasta.phy")
 

### 2017-03-04 all_gene_analysis/TextS1_all_gene_analysis.Rmd
renamed files

---
title: "All Gene Supplementary Material"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

###Load Libraries
```{r}
#libraries used for analysis
library(edgeR)
library(limma)
library(quadprog)
library(ruv)

#libraries required for graphics
library(ggplot2)
library(reshape2)
library(EDASeq)
library(RColorBrewer)
library(VennDiagram)
library(data.table)
library(dplyr)
library(stringr)
library(knitr)
library(pheatmap)

wd = "./"
colors <- c("#0571b0","#ca0020")
```

###Load our data
We first produce a matrix of read counts after aligning to both the Human and Plasmodium falciparum 3D7 reference genome.
Subread's featureCounts program is then used to allocate reads to gene features using the gff file PlasmoDB-12.0_Pfalciparum3D7.gff from PlasmoDB. 
For both Subread programs the default options are used.
```{r}
setwd(wd)
# # Next subreads featureCounts was used to allocate reads to genes. This was saved for re-use
# files <- Sys.glob("/home/users/allstaff/tonkin-hill.g/all_gene_analysis/alignment/*.sam")
# fc_genes_aligned_w_Pfhuman_withoutVivax <- featureCounts(files, annot.ext = "./data/PlasmoDB-12.0_Pfalciparum3D7.gff"
#                          , isPairedEnd=TRUE, isGTFAnnotationFile=TRUE, GTF.attrType="ID", GTF.featureType="gene"
#                          , nthreads=15)

#Load gene ID mappings which we use later on.
geneID_mappings <- fread("./data/geneID_mappings.txt", data.table = FALSE, stringsAsFactors = FALSE)
trim <- function (x) gsub("^\\s+|\\s+$", "", x)
s <- str_split(as.character(geneID_mappings$`[Previous ID(s)]`), ',')
geneID_mappings <- data.frame(current=rep(geneID_mappings$`[Gene ID]`, sapply(s, FUN=length)), old=unlist(s))
geneID_mappings$old <- str_trim(geneID_mappings$old)

#Load read counts (if featureCounts has been run previously)
load("./data/fc_genes_with_all_samples_aligned_human_Pf_WithOutVivax_subread_uH.RData")
fc2 <- fc_genes_aligned_w_Pfhuman_withoutVivax
t <- fc2$counts

#Rename samples to something more sensible
# colnames(t) <- gsub("alignment\\.sam","",colnames(t))
# colnames(t) <- gsub(".*vivax\\.","",colnames(t))
# colnames(t) <- gsub("_S.*","", colnames(t))
fc2$counts <- t

x <- DGEList(counts=fc2$counts, genes=fc2$annotation)
```

###Initial filter
We next filter out genes that do not have sufficient coverage across our samples for reliable estimates to be made. 
```{r}
setwd(wd)
before <- nrow(x$counts)
keep <- rowSums(cpm(x)>2) >= 10
x <- x[keep,,keep.lib.sizes=FALSE]
```
We require that there are at least 2 reads per million for at least 10 samples in order for a gene to be included in the analysis. This reduced the number of genes being considered from `r before` to `r nrow(x$counts)`.

###Remove low depth samples
```{r}
setwd(wd)
bplot <- melt(colSums(x$counts))
colnames(bplot) <- c("Read.Counts")
bplot$Sample <- rownames(bplot)

gg <- ggplot(bplot, aes(x=factor(Sample), y=Read.Counts)) + geom_bar(stat = "identity")
gg <- gg + theme_bw() + scale_y_sqrt(breaks=c(0,10000,1000000,10000000,100000000))
gg <- gg + theme(axis.text.x = element_text(size=12, angle = 90)
                 , axis.text.y = element_text(size=12, angle = 0)
                 , axis.title=element_text(size=14,face="bold"))
gg <- gg + labs(x='Sample', y='Read Count') 
gg <- gg + geom_hline(aes(yintercept=1000000), col="red")

x$counts <- x$counts[,!grepl("SFD1|SFM9|SFC\\.025", colnames(x$counts))]
filt.annnotation <- fc2$annotation[fc2$annotation$GeneID %in% rownames(x$counts),]
dge <- DGEList(counts=x$counts, genes=filt.annnotation)

gg
```
Samples SFD1, SFM9 and SFC.025 were not sequenced to a sufficient depth to be reliable and thus are removed.

###Remove drug treated samples
Samples SFU2 and SFU3 were drug treated and are removed.
```{r}
x$counts <- x$counts[,!grepl("SFU2|SFU\\.3", colnames(x$counts))]
filt.annnotation <- fc2$annotation[fc2$annotation$GeneID %in% rownames(x$counts),]
dge <- DGEList(counts=x$counts, genes=filt.annnotation)

```

Now we take the log transform of our sample counts (RPKM) to estimate the proportion of different stages present.
```{r}
setwd(wd)
our_log_rpkm <- rpkm(dge)
our_log_rpkm <- log2(1+our_log_rpkm)
```

##Estimate Contributions from Different Life-cycle Stages

###Load Data
To estimate the proportion of different life-cycle stages in each of our samples we require a reference data set. Su et al provide such a reference. They performed RNA-seq at two gametocyte stages, an ookinete stage and four time points of erythrocytic stages representing ring, early trophozoite, late trophozoite, and schizont. As we do not expect any ookinete stage parasites we discarded this reference. The data was downloaded from PlasmoDB and as such read counts were reported in RPKM. To handle outlying counts in a sensible manner a log2 transformation of the RPKM values was performed. This was also done for our gene counts. By next fitting a mixture model we can estimate the proportion of different parasite stages that make up our samples.
```{r}
setwd(wd)
su_rpkm <- read.csv("./data/ rpkm_7SexualAndAsexualLifeStages_suetal.csv", header=TRUE, sep=",")

#ignore the Ookinete profile as this should be absent from our data.
su_rpkm$Ookinete <- NULL
rownames(su_rpkm) <- su_rpkm$ID
su_rpkm$ID <- NULL

#take the log transform of the data which helps to manage outlier genes.
su_log_rpkm <- log2(1+su_rpkm)
```

###Mixture Model
Given a set of stages $S$ and a set of genes $g_i: i\in \{1,...,N\}$ we want to minimise

$$
\sum_{i=1}^N \left( g_{i,sample} -  \sum_{s \in S} \pi_s g_{i,s} \right)^2
$$

Subject to constraints

$$
\sum_{s\in S} \pi_s = 1
$$

$$
\pi_s > 0
$$

Here $\pi_s$ is proportion of stage $s$ that is attributed to having contributed to our sample. This model was fit using R's solve.QP function.

```{r}
setwd(wd)
findMix <- function(Y, X){  
 X[is.na(X)] <- t(replicate(ncol(X), apply(X,1,mean, na.rm=T)))[is.na(X)]
 Rinv <- solve(chol(t(X) %*% X))
 C <- cbind(rep(1,ncol(X)), diag(ncol(X)))
 b <- c(1,rep(0,ncol(X)))
 d <- t(Y) %*% X  
 qp <- solve.QP(Dmat = Rinv, factorized = TRUE, dvec = d, Amat = C, bvec = b, meq=1)
 sol <- qp$solution
 sol[sol<1e-10] <- 0
 return(sol)
}
```

###Calculate Sample Life-Cycle Profiles
We fit this mixture model to each of our samples in turn. The results are then plotted. A column represents each sample in the plot below. As we are fitting proportion parameters each columns values must add to 1 over the 5 stages. Note that we have also combined the estimated proportions for the gametocyte samples into one gametocyte variable. The columns are coloured by phenotype.
```{r, fig.height=7, fig.width=12}
setwd(wd)
#First get the intersection of the genes for the two datasets and order them by rowname.
inter <- intersect(rownames(our_log_rpkm), rownames(su_log_rpkm))

O <- our_log_rpkm[rownames(our_log_rpkm) %in% inter, ]
O <- O[order(rownames(O)),]
S <- su_log_rpkm[rownames(su_log_rpkm) %in% inter, ]
S <- S[order(rownames(S)), ]

#Now lets fit some samples!
ourPlotData <- data.frame()
for (i in 1:ncol(O)){
  mix <- findMix(O[,i], as.matrix(S))
  ourPlotData <- rbind(ourPlotData, data.frame(sample=rep(colnames(O)[i], ncol(S)), stage=colnames(S), proportion=mix))
}

#Organise the results.
ourPlotData$stage <- gsub("Gametocyte.*","Gametocyte",ourPlotData$stage)
ourPlotData <- aggregate(proportion~sample+stage,data=ourPlotData,FUN=sum)
ourPlotData <- within(ourPlotData, stage <- factor(stage, levels=c("Ring", "Early.Trophozoite", "Late.Trophozoite", "Schizont","Gametocyte")))
ourPlotData$phenotype <- ifelse(substring(ourPlotData$sample,1,1)=="I" ,"non-severe", "severe")

#Make a pretty plot.
gg <- ggplot(ourPlotData, aes(x=factor(sample), y=proportion, fill=factor(phenotype))) + geom_bar(stat="identity")
gg <- gg + scale_fill_manual(values = c("non-severe" = "#2c7bb6", "severe" = "#d7191c"))
gg <- gg + facet_wrap(~ stage, ncol = 1)
gg <- gg + theme_bw()
gg <- gg + theme(axis.text.x = element_text(size=12, angle = 90)
                 , axis.text.y = element_text(size=12, angle = 0)
                 , axis.title=element_text(size=14,face="bold")
                 , strip.text.x = element_text(size = 16, face="bold"))
gg <- gg + labs(x='Sample', y='Proportion', fill='Stage')
gg <- gg + theme(legend.text=element_text(size=14))
gg <- gg + theme(legend.key.size =  unit(0.25, "in"))
gg <- gg + theme(legend.title = element_text(size=16, face="bold"))
gg <- gg + guides(fill=guide_legend(title="Phenotype"))
gg
```

##Investigate Differential Expression
We now perform the Voom transformation on our count matrix. Prior to the transformation we use TMM normalisation to account for the different library sizes present in our sample. TMM was chosen as it is more robust to outliers. 
From the Voom variance trend it is evident that there are a number of genes that have both high variability and high expression.
After looking at some of the main offenders these were identified to be influenced by staging in the parasite. Consequently it is important to account for life-cycle stage when conducting the differential expression analysis.
```{r}
options(width=150)
setwd(wd)
categories<- as.factor(substring(rownames(dge$samples),1,1))
dge <- calcNormFactors(dge, method="TMM")
dge$samples$group <- categories
design <- model.matrix(~group, data=dge$samples)

v <- voom(dge, design=design, plot=TRUE)
```

###PCA Normalising for Library Size

The PCA plot below indicates that accounting for library size alone does not cope with the impact of life-cycle stages. Notably SFC21 is clearly an outlier and was identified in the mixture model as containing later stage parasites as well as gametocytes. IFM14 is also slightly separate in the PCA plot and was identified as having a proportion of Schizont stage parasite present.
```{r}
setwd(wd)
plotPCA(v$E, col=colors[categories], cex=0.8, isLog=TRUE)
```

Further the Relative Log Expression (RLE) plot below suggests that there is something different about SFC21 as well as highlighting IFM060 as being different.
```{r}
setwd(wd)
plotVoomRLE <- function(E, colours){
  mn <- apply(E, 1, median) 
  rle <- data.frame(sweep(E, MARGIN=1, STATS=mn, FUN='-')) 
  boxplot(rle,col=colours, outline=FALSE, las=2, ylim=c(-7,7))
  abline(h = 0, col = "black")
}
plotVoomRLE(v$E, colors[categories])
```

###PCA Normalising for Library Size, Ring, Gametocyte and Schizont effects.
To investigate the impact of life-cycle stages in our analysis we fit a linear model using limma's removeBatchEffect. Initially ring, gametocyte and schizont parameters were fit along with the disease phenotype factor. This was because for the most part the Trophozoite stages could be represented by accounting for these stages as they would be present implicitly in the model.

The PCA indicates that we have dealt with the outlying SFC21 sample as well as other less outlying samples such as IFM14 which showed a proportion of Schizont stage parasites in the mixture model. Unfortunately due to the low number of samples that have proportions of Schizont and Gametocytes estimating differential differential expression in severe disease becomes impossible. This is because there is insufficient data to estimate these stage parameters and consequently the uncertainty in the model is too high to find significant differences between severe and non-severe disease.

```{r}
setwd(wd)
covs <- data.frame(v$design[,2])
ourPlotData$phenotype <- NULL
c <-  dcast(ourPlotData, sample ~ ..., value.var="proportion")
covs <- merge(covs, c, by.x=0, by.y="sample")

colnames(covs) <- c("sample", "disease", colnames(c)[2:ncol(c)])
rownames(covs) <- covs$sample
covs$sample <- NULL
covs <- covs[match(colnames(v$E), rownames(covs)),]
stopifnot(colnames(v$E)==rownames(covs))
# head(covs)

covs <- covs[,c(1,2,5,6)]
# head(covs)

mod = model.matrix(
  as.formula(paste("~", paste(colnames(covs), collapse=" + "), sep="")), data=covs)

norm_counts_ring <- removeBatchEffect(v$E, covariates=mod[,3:ncol(mod)], design=mod[,1:2])
plotPCA(norm_counts_ring, col=colors[categories], cex=0.8, isLog=TRUE)
```

The RLE plot also suggests we have dealt with SFC21 however accounting for these stages does not seem to have accounted for the differences we see in the IFM060 sample. The mixture model did not identify IFM060 as being noticeably different from  the other samples and consequently it is likely that there is some other unknown factor impacting on its expression.
```{r}
setwd(wd)
plotVoomRLE(norm_counts_ring, col=colors[categories])
```

###PCA Normalising for Library Size and Ring Stage Effects

We then investigate the impact of only accounting for the Ring stage parameter in the model. This gives us more power with which to identify differential expression related to severe disease. However, from the PCA plot below it is clear that accounting for Ring stage alone is insufficient and fails to adequately deal with SFC21.
```{r}
setwd(wd)
covs <- covs[,c(1,2)]

mod = model.matrix(
  as.formula(paste("~", paste(colnames(covs), collapse=" + "), sep="")), data=covs)

norm_counts_ring <- removeBatchEffect(v$E, covariates=mod[,3:ncol(mod)], design=mod[,1:2])
plotPCA(norm_counts_ring, col=colors[categories], cex=0.8, isLog=TRUE)
```

We are however able to identify differentially expressed genes in this less parameterised model. The summary of results accounting only for the Ring stage variable are given below. This test was performed using the limma-Voom pipeline. Ideally we would be able to keep both the power in out tests and account for the remaining unwanted factors. To achieve this we make use of RUV-4 to estimate dummy variables that represent multiple unknown factors of unwanted variation.
```{r}
setwd(wd)
stopifnot(colnames(dge)==rownames(mod))
v2 <- voom(dge, design=mod, plot=FALSE)
fit <- lmFit(v2, mod)
fit <- eBayes(fit, robust=TRUE)

summary(de1<-decideTests(fit, adjust.method="BH", p.value=0.05))
```

###Estimate Unwanted Factors of Variation
To cope with the outlier samples such as SFC21 whilst retaining statistical power we estimate factors of unwanted variation using RUV-4.\
RUV-4 requires a set of control genes which are believed to be relatively unaffected by the condition of interest, in our case this is disease severity. RUV-4 is however fairly robust to the choice of control genes.\
We chose to use the 1000 genes found to have the highest p-values in the experiment of Vignali et al as controls. The experiment of Vignali et al looked at differences in expression between pregnancy associated malaria and childhood malaria. Although this is not the same association we are investigating it was considered close enough to perform adequately as a source of control genes. Further an analysis using SVA which does not require controls and rather estimates them empirically was found to produce similar results with only 5 out of 119 genes found using SVA missing from the RUV-4 approach. Further RUV-4 attempts to recover some variation that may be lost in the vanilla SVA approach when the factor of interest (severe disease) correlates with an unwanted factor. Consequently it leads to a higher number of differentially expressed genes being found. It was also thought that having an independently sourced set of control genes was preferable to empirically deriving such a set.

```{r}
setwd(wd)
contol_genes_vignali <- read.table("./data/contol_genes_vignali.txt", quote="\"")

ctrl_vignali <- geneID_mappings$current[geneID_mappings$old %in% contol_genes_vignali$V1]
length(ctrl_vignali)
empirical_controls <- ctrl_vignali[0:length(ctrl_vignali)]
empirical_controls <- rownames(v$E) %in% empirical_controls

categoriesRUV <- categories
categoriesRUV <- data.matrix(as.numeric(categoriesRUV=="S"))
ring <- data.matrix(mod[,c(1,3)])
genes <- data.matrix(t(v$E))

ruv <- RUV4(genes, categoriesRUV, empirical_controls, 3, Z=ring)

modRUV = cbind(mod,ruv$W)

norm_counts_ring_ruv <- removeBatchEffect(v$E, covariates=modRUV[,3:ncol(modRUV)], design=modRUV[,1:2])
```
Note that although we used 1000 genes from the experiment of Vignali et al only `r sum(empirical_controls)` were found in our reduced expression matrix.\
\
The PCA plot below indicates that we have dealt with most of the problems due to staging. Here we have included the Ring stage proportion from the mixture model as well as three factors of unwanted variation estimated using RUV-4. 
```{r}
setwd(wd)
plotPCA(norm_counts_ring_ruv, col=colors[categories], cex=0.8, isLog=TRUE)
```

The RLE plot also indicates that we have succesfully dealt with most of the staging issues. Both SFC21 and IFM14 have reasonable box plots. Moreover, IFM060 has a nicer boxplot indicating that in using RUV-4, not only have we dealt with staging issues, we have also accounted for some other unknown confounding factors that would have impacted our results. The variation in each sample has also been reduced which is indicated in the reduced height of the boxplot whiskers. Finally, the results were found to be relatively robust to the removal of the outlier samples such as SFC21 and IFM14. Consequently, this approach seems to succesfully deal with both life-cycle staging issues and other unknown impacting factors.

```{r}
setwd(wd)
plotVoomRLE(norm_counts_ring_ruv, colors[categories])
```

###Identify DE Genes using Limma and Voom
We now use the limma-voom pipeline, with the robust ebayes option which handles dispersion outliers. This was done to further ensure our results were less likely to be affected by outlier samples. Multiple testing correction was performed using the Benjamini-Hochberg approach.

```{r}
setwd(wd)
stopifnot(colnames(dge)==rownames(modRUV))
v2 <- voom(dge, design=modRUV, plot=F)
fit <- lmFit(v2, modRUV)
fit <- eBayes(fit, robust=TRUE)
s <- summary(de2<-decideTests(fit, adjust.method="BH", p.value=0.05))
s
```

Overall `r sum(s[c(1,3),2])` genes were found to be differentially expressed between severe and non-severe cases of malaria with a false dicovery rate threshold of 0.05.\
\
A good check for validity of our model it to look at the distribution of the resulting p-values. If everything is okay we would expect a uniform distribution except near $p=0$ where we would hope to see a spike. The barplot below indicates this is what we see which is helpful in affirming that we have done okay in removing unwanted variation without negatively impacting on the variation due to the variable of interest (diseas severity).
```{r}
setwd(wd)
top_ring_ruv <- topTable(fit, coef=2, p.value=0.1, sort.by="p", number=Inf, adjust.method="BH", confint=TRUE)
hist<- data.frame(topTable(fit, coef=2, number=Inf))
gg <- ggplot(hist, aes(x=P.Value)) + geom_histogram(binwidth = 0.01)
gg <- gg + theme_bw()
gg <- gg + theme(axis.text.x = element_text(size=12)
                 , axis.text.y = element_text(size=12, angle = 0)
                 , axis.title=element_text(size=12,face="bold"))
gg <- gg + labs(x='P value', y='Count')
gg
```

###Results table
Below is a table of genes that were found to be differentially expressed, ordered by their respective p-values. The 95% confidence intervals are given with the estimated LFC as well as the BH adjusted p-values and  the log-odds that the gene is differentially expressed. Here we report for interest sake all genes up to a adjusted p-value threshold of 0.1.
```{r kable}
setwd(wd)
#Load additional gene information
GeneAnnotationPlasmoDB <- fread("./data/GeneAnnotationPlasmoDB.txt"
                                , data.table = FALSE, na.strings = "N/A")

#First lets import some additional gene information to make a nicer table
GeneAnnotationPlasmoDB <- GeneAnnotationPlasmoDB[, colnames(GeneAnnotationPlasmoDB) %in% c("[Gene ID]", "[Product Description]",	"[Gene Name or Symbol]")]
top_ring_ruv <- merge(top_ring_ruv, GeneAnnotationPlasmoDB, by.x=0, by.y="[Gene ID]", all.x=TRUE)
top_ring_ruv <- top_ring_ruv[with(top_ring_ruv, order(adj.P.Val)), ]

#Now print the table
top_ring_ruv$P.Value <- format(top_ring_ruv$P.Value, scientific=TRUE, digits=3)
top_ring_ruv$adj.P.Val <- format(top_ring_ruv$adj.P.Val, scientific=TRUE, digits=3)
kable(top_ring_ruv[,c(2:4,6:ncol(top_ring_ruv))], digits=3)
```

##Heatmap of deregulated genes
```{r}
col_annot <- data.frame(row.names=colnames(v$E), stringsAsFactors = F)
col_annot$phenotype[substring(colnames(v$E),1,1)=="I"] <- "non-severe"
col_annot$phenotype[substring(colnames(v$E),1,1)=="S"] <- "severe"
col_annot$phenotype <- as.factor(col_annot$phenotype)

pheatmap(v2$E[rownames(v$E) %in% top_ring_ruv$GeneID,]
        , show_rownames = FALSE
        , scale = "row"
        , annotation_col = col_annot
        , annotation_colors = list(phenotype=c(`non-severe`=colors[1], `severe`=colors[2])))
```

##Gene sets from Daily et al and Shock et al

We now test using roast and camera for deregulation of gene sets obtained from the papers:

Daily, J. P., D. Scanfeld, N. Pochet, K. Le Roch, D. Plouffe, M. Kamal, O. Sarr, et al. 2007. “Distinct Physiological States of Plasmodium Falciparum in Malaria-Infected Patients.” Nature 450 (7172): 1091–95.

Shock, Jennifer L., Kael F. Fischer, and Joseph L. DeRisi. 2007. “Whole-Genome Analysis of mRNA Decay in Plasmodium Falciparum Reveals a Global Lengthening of mRNA Half-Life during the Intra-Erythrocytic Development Cycle.” Genome Biology 8 (7): R134.

```{r}
setwd(wd)

#Load gene lists from Daily et al and Shock et al
daily_figure2 <- fread("./data/daily_expression_figure2.csv"
                          , data.table = FALSE, header=TRUE)

JLshock <- c("PF3D7_0320800","PF3D7_0410400","PF3D7_0507600","PF3D7_0520300","PF3D7_0720000","PF3D7_0811300","PF3D7_0819900","PF3D7_0909400","PF3D7_0909900","PF3D7_1006100","PF3D7_1032100","PF3D7_1103800","PF3D7_1106300","PF3D7_1107000","PF3D7_1124400","PF3D7_1128600","PF3D7_1209200","PF3D7_1224300","PF3D7_1235300","PF3D7_1307000","PF3D7_1308900","PF3D7_1325000","PF3D7_1340100","PF3D7_1364500","PF3D7_1427800","PF3D7_1443300","PF3D7_1443500","PF3D7_1449700","PF3D7_0519500")

#Load entrex geneID mappings
entrez_geneID_mappings <- fread("./data/geneID_mappings.txt", data.table = FALSE, stringsAsFactors = FALSE)

upcluster2 <- daily_figure2$Gene[daily_figure2$`Up regulated in Cluster 2?`=="Y"]
upcluster1 <- daily_figure2$Gene[daily_figure2$`Up regulated in Cluster 2?`=="N"]
glycolisis <- daily_figure2$Gene[daily_figure2$type=="Glycolisis"]
tricarboxylic <- daily_figure2$Gene[daily_figure2$type=="Tricarboxylic acid cycle"]
fatty_acid <- daily_figure2$Gene[daily_figure2$type=="Fatty acid metabolism"]

upcluster2 <- as.vector(geneID_mappings$current[geneID_mappings$old %in% upcluster2])
upcluster1 <- as.vector(geneID_mappings$current[geneID_mappings$old %in% upcluster1])
glycolisis <- as.vector(geneID_mappings$current[geneID_mappings$old %in% glycolisis])
tricarboxylic <- as.vector(geneID_mappings$current[geneID_mappings$old %in% tricarboxylic])
fatty_acid <- as.vector(geneID_mappings$current[geneID_mappings$old %in% fatty_acid])

indices <- ids2indices(list(up_in_starvation=upcluster1, up_int_vitro=upcluster2, glycolisis=glycolisis, tricarboxylic=tricarboxylic, fatty_acid=fatty_acid, mRNA_degradation=JLshock), rownames(v2$E))
mroast(v2, index=indices, design=modRUV, contrast=2, nrot=10000, set.statistic="floormean")
camera(v2, index=indices, design=modRUV, contrast=2)
```

##KEGG analysis
```{r}
# setwd(wd)
# 
# entrez_de <- entrez_geneID_mappings$`[Entrez Gene ID]`[entrez_geneID_mappings$`[Gene ID]` %in% top_ring_ruv$GeneID]
# entrez_de <- entrez_de[entrez_de!="null"]
# entrez_de <- str_split_fixed(entrez_de, pattern = ",", n = 2)[,1]
# 
# kegg <- kegga(entrez_de, species.KEGG="pfa", convert=TRUE)
# kable(topKEGG(kegg, n=Inf))
```

##GO analysis
```{r}
# go <- goana(entrez_de, species.KEGG="pfa", convert=TRUE)
# kable(topGO(go, n=100))
```

####System Information
```{r}
setwd(wd)
sessionInfo()
```

### 2017-03-04 segment_level/TextS6_segment_level_analysis.Rmd
renamed files

---
title: "Segment Level Analysis"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

##Load libraries

```{r}
library(DESeq2)
library(ggplot2)
library(edgeR)
library(Rsubread)
library(data.table)
library(RFLPtools)
library(EDASeq)
library(igraph)
library(knitr)
library(pheatmap)
library(plyr)
library(dplyr)
library(stringr)
library(dynamicTreeCut)
library(moduleColor)

colors <- c("#0571b0","#ca0020")
wd <- "./"
```

##Homology blocks of Rask et al

Reads are aligned using Subread v1.4.6, mapping each read to its best match from a combined reference of the seperate sample denovo assembled contigs, the human reference (GRCh38) and the Pfalciparum reference genome (plasmoDB-11.1_Pfalciparum3D7_Genome) with the VAR gene regions masked out. 
```
subread-buildindex -o ref_index GRCh38.fasta plasmoDB-11.1_Pfalciparum3D7_Genome.fasta seperate_denovo_VAR.fasta
subread-align -i ref_index -r read_1.fastq -R read_2.fastq -o output.sam 
```

The homology blocks of Rask et al were identified in the translated seperate sample VAR contigs using HMMER v3.1b1. A score threshold of 9.7 was chosen as was used in Rask et al.

Note: homology blocks 37, 214, 300, 330, 337, 362, 485, 511, 528, 548, 552, 566, 598, 627 and 628 were not succesfully downloaded from http://www.cbs.dtu.dk/services/VarDom/ and were excluded from the analysis.

```
hmmsearch --nonull --noali --max --incdomT 9.97 --domT 9.97 --domtblout output.txt rask_hmm_file.hmm translated_seperate_denovo_VAR.fasta
```

The locations of these hits were then transformed to their respective locations in the original contigs fasta file and a SAF annotation file was generated. RfeatureCounts was then used to obtain read counts for each of these hits. Counts from hits to the same homology block were then combined.

```{r}
setwd(wd)
# files <- Sys.glob("/home/users/allstaff/tonkin-hill.g/gene_level_var_analysis/alignments_seperate_assemblies_best1/*.sam")
# block_counts <- featureCounts(files, annot.ext = "./data/rask_blocks/rask_blocks_in_seperate_assembly.SAF"
#                             , isPairedEnd=TRUE, isGTFAnnotationFile=FALSE
#                             , countChimericFragments=TRUE
#                             , allowMultiOverlap=TRUE
#                             , useMetaFeatures=FALSE
#                             , nthreads=20)
# 
# #combine counts for different runs of the same sample
# colnames(block_counts$counts) <- gsub(".*_best1\\.","",colnames(block_counts$counts))
# colnames(block_counts$counts) <- gsub("_S.*","",colnames(block_counts$counts))
# temp <- block_counts
# temp$counts <- t(rowsum(t(temp$counts), group = rownames(t(temp$counts))))
# dim(temp$counts)
# rask_block_counts <- temp
# 
# saveRDS(rask_block_counts, "./data/rask_block_counts.rds")

rask_block_counts <- readRDS("./data/rask_block_counts.rds")
```

Again we remove samples SFD1, SFM9 and SFC.025 as they were not sequenced to a sufficient depth.
```{r}
rask_block_counts$counts <- rask_block_counts$counts[,!(colnames(rask_block_counts$counts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
```

##Investigate block classifications of Rask et al
Combine counts based on homology block annotations.
```{r}
prev_block_counts <- data.frame(rask_block_counts$counts)
prev_block_counts$blockType <- unlist(lapply(rownames(prev_block_counts), function(x){
  temp <- unlist(strsplit(x, "_"))
  return(paste("block",temp[length(temp)-1],sep="_"))
  }))
prev_block_counts <- ddply(prev_block_counts, "blockType", numcolwise(sum))
rownames(prev_block_counts) <- prev_block_counts$blockType
prev_block_counts$blockType <- NULL
prev_block_counts <- as.matrix(prev_block_counts)

x_blocks <- DGEList(counts=prev_block_counts)
x_blocks <- calcNormFactors(x_blocks, method="RLE")
categories <- as.factor(substring(rownames(x_blocks$samples),1,1))
```
A total of `r dim(x_blocks)[1]` blocks from the those of Rask et al are identified in the samples.

##PCA plot for homology blocks

Below is a PCA plot from the normalised counts for the homology blocks of Rask et al.

```{r}
plotPCA(cpm(x_blocks,normalized.lib.sizes=TRUE, log=T, prior.count=0.5), col=colors[categories], cex=0.8, isLog=TRUE)
title("VAR PCA - Rask Homology Blocks")
```

##Investigate differential expression

Differential expression analysis was conducted with the default settings for DESeq2.

```{r}
colData <- data.frame(disease=categories)
rownames(colData) <- colnames(x_blocks$counts)
dds_block <- DESeqDataSetFromMatrix(countData = x_blocks$counts
                              , colData = colData
                              , design = ~disease)
dds_block <- estimateSizeFactors(dds_block)
dds_block <- DESeq(dds_block) 
res <- results(dds_block,contrast=c("disease","S","I"), pAdjustMethod = "BH")
summary(res, alpha=0.05)
```

In total `r sum(res$padj[!is.na(res$padj)]<0.05)` blocks were found to be differentially expressed, with `r sum(res$padj[!is.na(res$padj) & (res$log2FoldChange>0)]<0.05)` upregulated in severe disease.

##A heatmap of the Up-regulated homology blocks

To investigate the expression of homology blocks at a higher resolution, we can look at a heatmap of expression. Here samples are clustered by average linkage hierarchical clustering.

```{r}
res_blocks <- res
resOrdered <- res_blocks[order(res_blocks$padj),]
resSig <- subset(resOrdered, padj < 0.05)

resSigLFC <- data.frame(subset(resSig, log2FoldChange > 0))

select <- order(rowMeans(counts(dds_block, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds_block)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resSigLFC),]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe" 
annotData$phenotype[annotData$disease=="S"] <- "severe"
annotData$disease <- NULL
pheatmap(log2.norm.counts, 
         cluster_rows=TRUE, show_rownames=TRUE, 
         cluster_cols=TRUE, annotation_col=annotData, 
         annotation_colors = anno_colors,
         clustering_method = "average")
```

#Novel segment investigation

To define novel segments we aligned each of the major domain classes identified using gismo v2.0. The alignments were then segmented by defining conserved regions where there were 7 or more consecutive columns with greater than 95% occupancy. A seperate fasta file was constructed for each conserved region and each region inbetween the conserved regions. Gap positions were removed from these protein segments and each segment was kept if it was at least 7aa in length.

The resulting position of these protein segements was then translated into a position in the original contigs file and a SAF annotaion file was created.

Hierarchical clustering was performed on each of the seperated protein fasta files using CD-HIT at levels 0.97,0.95,0.9,0.85,0.8,0.75,0.7,0.65,0.6 and 0.5. Global pairwise identity was required by running CD-HIT as
```
cd-hit -n 2 -d 1000 -l 2 -G 0 -aS 1 -aL 1 -c identity_threshold -i input_fasta -o output -g 1 -AL 99999999
```
A python script available in the github repository implements this stage.

By ensuring that segments are only clustered with those from the same region on each of the domain classes we are better able to infer the relationships between these segments and the domains. Furthermore, by investigating each of the regions at different clustering resolutions we are better able to distinguish those segements that may relate to severe disease.

##Assign read counts to segements

Read counts are assigned to segments in the same manner as for the block annotations of Rask et al.

```{r}
setwd(wd)
# files <- Sys.glob("/home/users/allstaff/tonkin-hill.g/gene_level_var_analysis/alignments_seperate_assemblies_best1/*.sam")
# block_counts <- featureCounts(files, annot.ext = "./combined_blocks.SAF"
#                             , isPairedEnd=TRUE, isGTFAnnotationFile=FALSE
#                             , countChimericFragments=TRUE
#                             , allowMultiOverlap=TRUE
#                             , useMetaFeatures=FALSE
#                             , nthreads=20)
# 
# #combine counts for different runs of the same sample
# colnames(block_counts$counts) <- gsub(".*_best1\\.","",colnames(block_counts$counts))
# colnames(block_counts$counts) <- gsub("_S.*","",colnames(block_counts$counts))
# temp <- block_counts
# temp$counts <- t(rowsum(t(temp$counts), group = rownames(t(temp$counts))))
# dim(temp$counts)
# gismo_block_counts <- temp
# 
# saveRDS(gismo_block_counts, "./gismo_block_counts.rds")
gismo_block_counts <- readRDS("./data/gismo_block_counts.rds")
```

##Remove low coverage samples

Again we remove samples SFD1, SFM9 and SFC.025 as they were not sequenced to a sufficient depth.

```{r}
gismo_block_counts <- gismo_block_counts$counts[,!(colnames(gismo_block_counts$counts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
```

##Summarise at different clustering heights

Here we merge the clustering obtained using cd-hit with the count information obtained for each segment.

```{r}
clusters <- fread("./data/combined_blocks_clustered.csv", data.table = FALSE, header = TRUE)
stopifnot(nrow(gismo_block_counts)==nrow(clusters))
gismo_block_clusters <- merge(gismo_block_counts, clusters, by.x=0 , by.y="Transcript")
```

##Create hierarchical structure

Starting with no clustering, individual domain counts are combined at the different hierarchical clustering levels.

```{r}
#first create an edgelist
edges <- list()

#Add in root node to edge list
edges <- rbind(edges,cbind("root", paste(as.character(clusters[,2]), colnames(clusters)[2], sep="_")))

#Add in remaining edges at successive levels of the hierarhcy
for (j in 3:(ncol(clusters))){
      edges <- rbind(edges,cbind(paste(as.character(clusters[,j-1]), colnames(clusters)[j-1], sep="_"), paste(as.character(clusters[,j]), colnames(clusters)[j], sep="_")))
}

#Remove duplicate edges caused by the similar sequences propogating up the tree together.
edges <- edges[!duplicated(edges),]
edges <- apply(as.matrix(edges), 2,as.character)

#now sum counts at the different clustering heights
results <- data.frame()
pca.df <- data.frame()
all_clusters.matrix <- data.frame()
for (c in colnames(clusters)[2:(ncol(clusters))]) {
  #First sum based upon cluster at level c
  gismo_block_clusters.dt <- data.table(gismo_block_clusters)
  gismo_block_clusters.dt <- gismo_block_clusters.dt[,lapply(.SD, sum, na.rm=TRUE)
                                                       , by=c
                                                       , .SDcols=colnames(gismo_block_clusters)[2:42]]

  #Add summed counts to the all_clusters matrix
  gismo_block_clusters.matrix <- as.matrix(gismo_block_clusters.dt)
  gismo_block_clusters.matrix <- gismo_block_clusters.matrix[,2:ncol(gismo_block_clusters.matrix)]
  rownames(gismo_block_clusters.matrix) <- paste(gismo_block_clusters.dt[[1]], c, sep="_")

  #check columns are the same after first
  if (c!=colnames(clusters)[[2]]){
    stopifnot(colnames(all_clusters.matrix)==colnames(gismo_block_clusters.matrix))
  }

  all_clusters.matrix <- rbind(all_clusters.matrix, gismo_block_clusters.matrix)

  #generate points for PCA
  dgePCA <- DGEList(counts=gismo_block_clusters.matrix)
  colData <- data.frame(disease=as.factor(substring(colnames(gismo_block_clusters.matrix),1,1)))
  rownames(colData) <- rownames(dgePCA$samples)
  ddsPCA <- DESeqDataSetFromMatrix(countData = dgePCA$counts#dom_counts_clusters.matrix
                            , colData = colData
                            , design = ~disease)
  cts <- counts(ddsPCA)

  ddsPCA <- estimateSizeFactors(ddsPCA)
  pca <- plotPCA(varianceStabilizingTransformation(ddsPCA), intgroup=c("disease"), returnData = TRUE)
  pca$level <- c
  pca.df <- rbind(pca.df, pca)
}
```

##PCA plots at different levels

Below PCA plots are shown for each of the different clustering levels based off of the normalised counts for that level. It inidicates that the severe samples are most tightly clustered at the 50% identity level.

```{r}
pca.df$phenotype[pca.df$disease=='S'] <- "severe"
pca.df$phenotype[pca.df$disease=='I'] <- "non-severe"
gg <- ggplot(pca.df, aes(PC1, PC2, color=phenotype)) + scale_color_manual(values=colors)
gg <- gg + geom_point()
gg <- gg + facet_wrap(~level)
gg <- gg + theme_bw() 
gg
```

##PCA plot for segment clusters at 50% identity
```{r}
#group counts at the 50% ID level
gismo_block_clusters.dt <- data.table(gismo_block_clusters)
gismo_block_clusters.dt <- gismo_block_clusters.dt[,lapply(.SD, sum, na.rm=TRUE)
                                                       , by='0.5'
                                                       , .SDcols=colnames(gismo_block_clusters)[2:42]]

gismo_block_clusters.matrix <- as.matrix(gismo_block_clusters.dt)
gismo_block_clusters.matrix <- gismo_block_clusters.matrix[,2:ncol(gismo_block_clusters.matrix)]
rownames(gismo_block_clusters.matrix) <- paste(gismo_block_clusters.dt[[1]], c, sep="_")

#generate points for PCA
my_domains <- DGEList(counts=gismo_block_clusters.matrix)
my_domains <- calcNormFactors(my_domains, method="RLE")
plotPCA(cpm(my_domains,normalized.lib.sizes=TRUE, log=T, prior.count=0.5), col=colors[categories], cex=0.8, isLog=TRUE)
title("VAR PCA - Segments clustered at 50% identity")
```

##Differential expression testing

First, to speed up computation and focus on only the most interesting domain clusters we remove clusters with less than 1 read per million which appear in less than 5 samples.

```{r}
dge_clust <- DGEList(counts=as.matrix(all_clusters.matrix))

#create design matrix
colData <- data.frame(disease=as.factor(substring(rownames(dge_clust$samples),1,1)))
rownames(colData) <- rownames(dge_clust$samples)

cpm(10, mean(dge_clust$samples$lib.size))

keep <- rowSums(cpm(dge_clust)>10) >= 5
dim(dge_clust)
dge_clust <- dge_clust[keep,,keep.lib.sizes=FALSE]
dim(dge_clust)
```

DESeq2 is now run with default parameters using the Benjamini-Yekutieli procedure for multiple testing correction.

```{r, cache=FALSE}
setwd(wd)

#create DESeq2 dataset
stopifnot(colnames(dge_clust)==rownames(colData))
dds <- DESeqDataSetFromMatrix(countData = dge_clust$counts
                            , colData = colData
                            , design = ~disease)

#estimate size factors
dds <- estimateSizeFactors(dds)

#run DESeq2 pipeline
dds <- DESeq(dds)

results <- results(dds, contrast=c("disease","S","I"), pAdjustMethod = "BY")
summary(results, alpha=0.05)
```

Match up p-values with edges in the tree structure.

```{r}
#Only look at significant nodes after multiple testing adjustment
Sig <- results[!is.na(results$padj),]
Sig <- Sig[Sig$padj<0.05,]

#Create tree from edge list.
tree <- graph.edgelist(edges)

#If nodes aren't tested (i.e didn't pass the conservartion filter) set there significane to NA
naNodes <- V(tree)$name[!(V(tree)$name %in% rownames(results))]
na1s = rep(1, length(naNodes))

#Create vectors of p-value with and without multiple testing adjustment.
unadjp <- c(na1s, results$pvalue)
adjp <- c(rep(NA, length(naNodes)), results$padj)
names(unadjp) <- c(naNodes, rownames(results))
names(adjp) <- c(naNodes, rownames(results))

#Check we've accounted for all nodes in tree
if (!all(names(unadjp) %in% unique(as.vector(edges)))) {
        stop("Names of elements in unadjp do not match names of tree nodes")
    }
if (!all(unique(as.vector(edges)) %in% names(unadjp))) {
        stop("Names of elements in unadjp do not match names of tree nodes")
    }

p.vals <- data.frame(unadjp, adjp = adjp)
rownames(p.vals) <- names(unadjp)

cluster_names <- clusters
for (i in 2:ncol(clusters)){
  cluster_names[,i] <- paste(as.character(cluster_names[,i]), colnames(clusters)[i], sep="_")
}

#Clean up segment names from transcript
rnames <- gsub(".*_DBL", "DBL", clusters$Transcript)
rnames <- gsub(".*_CIDR", "CIDR", rnames)
rnames <- gsub(".*_NTS", "NTS", rnames)
rnames <- gsub("_domNum.*_", "_", rnames)
rnames <- gsub("[0-9]*_", "_", rnames)

#Rename clusters with major domain class and location within domain
cluster_names <- apply(cluster_names, 2, paste, rnames, sep="_")
```

##Functions to explore the tree

These function return the ancestors and children of a given node in the hierarchical tree.

```{r}
#return the children of a node
get_children <- function(node, edges) {
    if (is.null(node)){
      return(NULL)
    }
    tree.el.tmp <- data.frame(parent = edges[, 1], child = edges[, 2], stringsAsFactors = F)
    children <- tree.el.tmp[which(tree.el.tmp$parent == node), "child"]
    children <- unlist(c(children, lapply(children, get_children, edges)))
    return(children)
}

#return the anscestors of a node
get_parents <- function(node, edges){
  if (is.null(node)){
      return(NULL)
    }
    tree.el.tmp <- data.frame(parent = edges[, 1], child = edges[, 2], stringsAsFactors = F)
    parent <- tree.el.tmp[which(tree.el.tmp$child == node), "parent"]
    parent <- unlist(c(parent, lapply(parent, get_parents, edges)))
    return(parent)
}

```

##Get significant nodes in trees

From those nodes that are found to be significant after multiple testing correction we keep only the most signifcant from one path through the tree. That is, if a node has an ancestor or child that has a smaller p-value, it will not be considered. This is achieved by iteratively selecting the most significant node and removing it's ancestors and children from the list of signifcant nodes.

```{r}
tempSig <- Sig
tempSig$cluster <- as.numeric(gsub(".*_","",rownames(tempSig)))
tempSig <- tempSig[order(tempSig$cluster),]
SigFilteredResults <- data.frame()
while (nrow(tempSig)>0){
  #Get most significant and add to final results
  trow <- tempSig[which.min(tempSig$padj),]
  SigFilteredResults <- rbind(SigFilteredResults, trow)

  #Find parents and child nodes of the current most significant node
  t <- rownames(trow)
  tChildren <- get_children(t, edges)
  tParents <- get_parents(t, edges)

  #Remove these nodes from those available for selection
  tempSig <- tempSig[!(rownames(tempSig) %in% c(t, tChildren, tParents)),]
  tempSig <- tempSig[order(tempSig$cluster),]
}

#Split into up and downregulated nodes
resUp <- SigFilteredResults[SigFilteredResults$log2FoldChange>0,]
resDown <- SigFilteredResults[SigFilteredResults$log2FoldChange<0,]
```

Overall we have found `r nrow(SigFilteredResults)` significantly DE clusters of which `r sum(SigFilteredResults$log2FoldChange>0)` are up-regulated.

Update final results to include information on the transcripts that contain them
```{r}
#Collect sequence labels for each cluster
final <- data.frame(SigFilteredResults)
sig_clusters <- as.numeric(gsub("_.*","", rownames(SigFilteredResults)))
levels <- gsub(".*_","", rownames(SigFilteredResults))
significantClusters <- list()
for (i in 1:length(levels)){
  significantClusters[[i]] <- clusters[clusters[,colnames(clusters)==levels[i]]==sig_clusters[i],]$Transcript
}
final$sequences <- I(significantClusters)

getBlock <- function(x){
  x <- x[[1]][1]
  x <- toString(x)
  x <- gsub(".*_DBL", "DBL", x)
  x <- gsub(".*_CIDR", "CIDR", x)
  x <- gsub(".*_NTS", "NTS", x)
  x <- gsub("_domNum.*_", "_", x)
  x <- gsub("[0-9]*_", "_", x)
  x <- gsub("[0-9]\\._", "_", x)
  return(x)
}
cluster_block_type <- unlist(lapply(final$sequences, getBlock))
names(cluster_block_type) <- rownames(final)
```

##Clustering significant domains based on co-expression
```{r}
# expression <- cpm(dge_clust, normalized.lib.sizes=TRUE, log=TRUE)[rownames(dge_clust) %in% rownames(resUp),]
select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
expression <- assay(nt)[rownames(nt) %in% rownames(resUp),]

dissim = dist(expression)
dendro = hclust(dissim, method = "complete")

AutoColor = NULL
for (deepSplit in 0:3)
  AutoColor = cbind(AutoColor, labels2colors(cutreeDynamic(dendro = dendro,
                    minClusterSize = 2,
                    cutHeight = NULL, method = "hybrid", deepSplit = deepSplit,
                    distM = as.matrix(dissim))))
AutoLabels = paste("Hybrid 'auto': dS =", c(0:3))

cut <- cutreeDynamic(dendro = dendro,
                    minClusterSize = 2,
                    cutHeight = NULL, method = "hybrid", deepSplit = 3,
                    distM = as.matrix(dissim))

rowLabels <- dendro$labels
# rowLabels <- rowLabels[dendro$order]
rowsAnnot <- data.frame(group=labels2colors(cut))
rownames(rowsAnnot) <- rowLabels

par(mfrow=c(2,1))
par(cex = 1.4);
par(mar = c(0,8.5,2,0));
plot(dendro,labels=F,main="Co-expression clustering with dynamic tree cut", sub="", xlab="")
par(mar = c(1,8.5,0,0));
plotHclustColors(dendro, AutoColor, 
            AutoLabels, 
            main = "")
```

##Heatmap of up-regulated clusters

```{r}
select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe"
annotData$phenotype[annotData$disease=="S"] <- "severe"
annotData$disease <- NULL

log2.norm.counts <- log2.norm.counts[dendro$order,]
pheatmap(log2.norm.counts, cluster_rows=FALSE, show_rownames=TRUE, cluster_cols=TRUE
         , annotation_col=annotData, annotation_colors = anno_colors
         , clustering_method = "average", cutree_rows=9
         , labels_row=paste(rownames(log2.norm.counts), cluster_block_type[rownames(log2.norm.counts)] ,sep=" - ")
         , annotation_row=rowsAnnot, fontsize_row=7)
```

##Heatmap of down-regulated clusters

```{r}
setwd(wd)

nt <- normTransform(dds)
expression <- assay(nt)[rownames(nt) %in% rownames(resDown),]
dissim = dist(expression)
dendro = hclust(dissim, method = "complete")
cut <- cutreeDynamic(dendro = dendro,
                    minClusterSize = 2,
                    cutHeight = NULL, method = "hybrid", deepSplit = 0,
                    distM = as.matrix(dissim))
rowLabels <- dendro$labels
rowsAnnot <- data.frame(group=labels2colors(cut))
rownames(rowsAnnot) <- rowLabels

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resDown),]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe"
annotData$phenotype[annotData$disease=="S"] <- "severe"
annotData$disease <- NULL

log2.norm.counts <- log2.norm.counts[dendro$order,]
pheatmap(log2.norm.counts, cluster_rows=FALSE, show_rownames=TRUE, cluster_cols=TRUE
         , annotation_col=annotData, annotation_colors = anno_colors
         , clustering_method = "complete", cutree_rows=10
         , labels_row=paste(rownames(log2.norm.counts), cluster_block_type[rownames(log2.norm.counts)] ,sep=" - ")
         , annotation_row=rowsAnnot, fontsize_row=6)
```

##Table of results
```{r}
setwd(wd)
# Now print the table
rownames(final) <- paste(rownames(final), cluster_block_type[rownames(final)] ,sep="-")
final$padj <- format(final$padj, scientific=TRUE, digits=3)
final$pvalue <- format(final$pvalue, scientific=TRUE, digits=3)
final$sequences <- unlist(lapply(final$sequences, paste, collapse=","))
kable(final, digits=3)
```

##Code for generating the tree plots
```{r}
setwd(wd)

write.table(results, file="./data/tree_figures/results.csv", sep=",", quote=F)
clust <- data.frame(clusters)
clust$transcript <- clust$Transcript
clust$Transcript <- NULL
write.table(clust, file="./data/tree_figures/segement_clusters_complete.csv", sep=",", quote=F, row.names=F)

get_root <- function(node, edges){
  parents = get_parents(node, edges)
  if (length(parents)==1){
    return(node)
  }
  else{
    return(parents[length(parents)-1])
  }
}

resSig <- results[!is.na(results$padj),]
resSig <- resSig[resSig$padj<=0.05,]

sig <- data.frame(clusters=rownames(SigFilteredResults), root=gsub("_.*", "", unlist(lapply(rownames(SigFilteredResults), get_root, edges))))
write.table(sig, file="./data/tree_figures/sig_clusters.csv", sep=",", row.names=FALSE, quote = FALSE)
```

=======
##System Information
```{r}
sessionInfo()
```

### 2017-03-04 transcript_level/TextS2_combined_assembly_transcript_analysis.Rmd
renamed files

---
title: "Combined assembly VAR Transcript DE Analysis"
author: "Gerry Tonkin-Hill"
date: "22/10/2015"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

###Load Libraries
```{r}
library(limma)
library(edgeR)
library(DESeq2)
library(genefilter)
library(EDASeq)
library(RColorBrewer)
library(VennDiagram)
library(knitr)
library(reshape2)

library(pheatmap)
library(dynamicTreeCut)
library(moduleColor)
library(ggplot2)

wd = "./"
colors <- c("#0571b0","#ca0020")
```

###Load Data
We load the read counts from the all gene analysis as well as the resulting design matrix. We also aligned each sample to a reference which included human, non-VAR falciparum, non-VAR vivax and our VAR gene transcripts from the combined sample assembly. Read counts that aligned to our transcripts were then retrieved. We use the previous all gene counts to achieve better estimates of the dispersion parameters in our model.
```{r}
setwd(wd)
load("./data/geneWithoutVivax_and_VARgene_countDB_subread_uH.RData")
var_gene_names <- read.delim("./data/var_gene_names.txt")
var_gene_names <- var_gene_names$X.Gene.ID.
#Load Non-var_genes
fc2 <- fc_genes_aligned_w_Pfhuman_withoutVivax
t <- fc2$counts

#Rename samples to something more sensible
# colnames(t) <- gsub("alignment\\.sam","",colnames(t))
# colnames(t) <- gsub(".*vivax\\.","",colnames(t))
# colnames(t) <- gsub("_S.*","", colnames(t))
nonVar <- t
nonVar.annot <- fc2$annotation

var <- var_counts$counts
var.annot <- var_counts$annotation
# colnames(var) <- gsub(".*alignments\\.","",colnames(var))
# colnames(var) <- gsub("alignment.sam","",colnames(var))
# colnames(var) <- gsub("_L00.*", "", colnames(var))
# colnames(var) <- gsub("_S.*", "", colnames(var))
var_names <- rownames(var)

stopifnot(colnames(var)==colnames(nonVar))
combined_counts <- rbind(nonVar, var)
combined_annot <- rbind(nonVar.annot, var.annot)
```

Again we remove samples SFD1, SFM9 and SFC.025 as they were not sequenced to a sufficient depth.
```{r}
combined_counts <- combined_counts[,!(colnames(combined_counts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
```

###Filter Low Count Genes
```{r}
stopifnot(rownames(combined_counts)==combined_annot$GeneID)
dge <- DGEList(counts=combined_counts, genes=combined_annot)
dge_combined <- dge

cpm(10, mean(dge_combined$samples$lib.size))

keep <- rowSums(cpm(dge)>1) >= 5
dge <- dge[keep,,keep.lib.sizes=FALSE]

categories <- as.factor(substring(colnames(combined_counts),1,1))

modRUV <- modRUV[match(colnames(dge), rownames(modRUV)),]
stopifnot(colnames(dge)==rownames(modRUV))
```
We filter out genes with low coverage. As many of the VAR gene transcripts are short in length and as there is still some redundancy in the transcripts (although small) the threshold for removal was reduced compared to the all gene analysis. Genes were kept if they had more than 0.5 reads per million in at least 10 samples. Before filtering there was a total of `r dim(dge_combined$counts)[1]` genes. After filtering `r dim(dge$counts)[1]` remained. Of the initial `r sum(rownames(dge_combined) %in% var_names)` VAR gene transcripts `r sum(rownames(dge) %in% var_names)` remained.

###All Gene PCA
To enable a comparison a plot of the all gene PCA is given. Note that this is slightly different to the one in the all gene analysis due to the less stringent filter.
```{r}
categories<- as.factor(substring(rownames(dge$samples),1,1))
dge <- calcNormFactors(dge, method="TMM")
dge$samples$group <- categories
design <- model.matrix(~group, data=dge$samples)

v <- voom(dge, design=design, plot=FALSE)

norm_counts_ring_ruv <- removeBatchEffect(v$E, covariates=modRUV[,3:ncol(modRUV)], design=modRUV[,1:2])
plotPCA(norm_counts_ring_ruv[!(rownames(norm_counts_ring_ruv) %in% c(var_gene_names,var_names)),], col=colors[categories], cex=0.8, isLog=TRUE)
title("Non-VAR PCA")
```

###VAR Transcript PCA (normalised for Ring and RUV-4 factors)
The remaining VAR gene transcripts are investigated in the following PCA plot. Interestingly we can see that although the non-severe samples are very spread out, the severe samples remain close together. This suggested that there is something in common in the VAR gene expression profiles of the severe samples. Exactly what produces this grouping is not clear from this PCA and further investigation is given later. However the grouping is consistent with the immune response we see in human populations where by people become immune to severe forms of malaria relatively quickly but take much longer to develop immunity to non-severe forms.
```{r}
plotPCA(norm_counts_ring_ruv[rownames(norm_counts_ring_ruv) %in% var_names,], col=colors[categories], cex=0.8, isLog=TRUE)
```

###VAR Transcript PCA (only normalised for library size)
Although using voom and RUV4 is robust to sources of unwanted variation it is conservative. Furthermore, VAR genes are generally expressed during the XXX stage and thus variation due to life-cycle is less of an issue when investigating VAR genes. To this effect the PCA plot below indicates the separation of the samples due to the VAR gene transcripts without accounting for life-cycle or the RUV-4 factors. Although the separation between the severe and non-severe samples is reduced the same grouping effect is apparent. By ignoring these factors we are also able to increase the sensitivity of our analysis. This potentially could increase the FDR, however it returns a number of genes that is appropriate for further biological analysis whilst including the majority of the genes found by the voom approach.
```{r}
plotPCA(cpm(dge,normalized.lib.sizes=TRUE, log=T, prior.count=1)[rownames(dge) %in% var_names,], col=colors[categories], cex=0.8, isLog=TRUE)
```

###Voom Analysis (including RUV-4 and Ring factors)
Initially we performed the analysis using the same approach as the all gene analysis. After running Voom the adjusted p-values for the VAR gene transcripts were then recalculated using the Benjamini-Hochberg approach, only taking into account the VAR transcripts.
```{r}
stopifnot(colnames(dge)==rownames(modRUV))
v2 <- voom(dge, design=modRUV, plot=F)
fit <- lmFit(v2, modRUV)
fit <- eBayes(fit)
deVoom <- decideTests(fit, adjust.method="BH", p.value=0.05)

top_ring_ruv <- topTable(fit, coef=2, p.value=1, sort.by="p", number=Inf, adjust.method="none", confint=TRUE)
top_ring_ruv_var <- top_ring_ruv[rownames(top_ring_ruv) %in% var_names,]
top_ring_ruv_var$adj.P.Val <- p.adjust(top_ring_ruv_var$adj.P.Val, method = "BH")
top_ring_ruv_var <- top_ring_ruv_var[top_ring_ruv_var$adj.P.Val<=0.05,]

voomUp <- rownames(top_ring_ruv_var[top_ring_ruv_var$logFC > 0,])
voomDown <- rownames(top_ring_ruv_var[top_ring_ruv_var$logFC < 0,])
voomUpDown <- c(voomUp, voomDown)
```
This identified `r length(voomUpDown)` VAR transcripts as being differentially expressed. Of these `r length(voomUp)` were upregulated in severe disease.

###DESeq2 Analysis (simplified design matrix)
In order to achieve a higher sensitivity we looked at the VAR gene transcripts using DESeq2 ignoring the Ring and RUV-4 factors. Note that we used the default DESeq2 pipeline for this analysis. This included using DESeq2 to filter out genes with low expression. It's default filter was less stringent to the one used in the voom analysis. Unlike the voom approach, only the VAR gene counts were used in this analysis. This reduced any dependancy that life-cycle might play on overall VAR expression and increased the sensitivity of the analysis. This approach is likely to have a higher FDR however in this case we are interested in identifying a larger number of VAR genes for further biological analysis.
```{r}
colData <- data.frame(disease=as.factor(substring(rownames(dge_combined$samples),1,1)))
rownames(colData) <- rownames(dge_combined$samples)
dds <- DESeqDataSetFromMatrix(countData = dge_combined$counts[rownames(dge_combined) %in% var_names,]
                              , colData = colData
                              , design = ~disease)
dds <- estimateSizeFactors(dds)
dds <- DESeq(dds) 
res <- results(dds,contrast=c("disease","S","I"))

resVar <- res[rownames(res) %in% var_names,]
simpleDE_resVar <- data.frame(resVar)
resOrdered <- resVar
resOrdered <- resOrdered[!is.na(resOrdered$padj),]
resOrdered <- resOrdered[resOrdered$padj<0.05,]
```

A summary of the DESeq2 analysis on the VAR gene transcripts is given below.
```{r}
summary(resVar, alpha=0.05)
```

Genes identified as significantly differentially expressed are then grouped according to LFC.
```{r}
de_varOnlyGenesUpDown <- rownames(resOrdered)

resUp <- resOrdered[resOrdered$log2FoldChange>0,]
resDown <- resOrdered[resOrdered$log2FoldChange<0,]
DESeqUp <- rownames(resUp)
DESeqDown <- rownames(resDown)
DESeqUpDown <- c(DESeqUp, DESeqDown)
```
Using this simplified approach a total of `r length(DESeqUpDown)` VAR transcripts were identified as differentially expressed. Of these `r length(DESeqUp)` were upregulated. This set of upregulated transcripts included the one identified by the voom analysis. Further, of the `r length(voomDown)` genes identified as down regulated by the voom analysis, `r length(intersect(voomDown, DESeqDown))` were found using DESeq2.

##Heatmap of up-regulated segment clusters

```{r}
setwd(wd)

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]
# log2.norm.counts <- log2.norm.counts[,colData$disease=="S"]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe" 
annotData$phenotype[annotData$disease=="S"] <- "severe"
# annotData <- annotData[annotData$disease=="S",]
annotData$disease <- NULL
pheatmap(log2.norm.counts, cluster_rows=TRUE, show_rownames=T, cluster_cols=TRUE, annotation_col=annotData, annotation_colors = anno_colors)
```

##Heatmap of down-regulated segment clusters

```{r}
setwd(wd)

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resDown),]
# log2.norm.counts <- log2.norm.counts[,colData$disease=="S"]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe" 
annotData$phenotype[annotData$disease=="S"] <- "severe"
# annotData <- annotData[annotData$disease=="S",]
annotData$disease <- NULL
pheatmap(log2.norm.counts, cluster_rows=TRUE, show_rownames=FALSE, cluster_cols=TRUE, annotation_col=annotData, annotation_colors = anno_colors)
```

#DESeq2 Transcript Results
```{r}
#Now print the table
resVar$pvalue <- format(resVar$pvalue, scientific=TRUE, digits=3)
resVar$padj <- format(resVar$padj, scientific=TRUE, digits=3)
kable(resVar, digits=3)
```

##Session Information
```{r}
sessionInfo()
```

### 2017-03-04 transcript_level/TextS3_seperate_assembly_CORSET_analysis.Rmd
renamed files

---
title: "Seperate assembly VAR Transcripts (Corset) DE Analysis"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

##Load libraries
```{r}
library(DESeq2)
library(ggplot2)
library(edgeR)
library(knitr)
library(pheatmap)
library(EDASeq)

wd <- "./"
colors <- c("#0571b0","#ca0020")
```

##Running CORSET
CORSET requires that reads are mapped to the transcripts allowing for multiple mappings of the same read. To achieve this bowtie verion 0.12.9 was run with options to allow for multiple mapping
```
bowtie-build comma_seperated_list_of_fasta_files bowtie_REF_VAR_seperate_assembly

bowtie -S --all --threads 30 bowtie_REF_VAR_seperate_assembly -1 read1.fq -2 read2.fq alignment_output.sam
```

CORSET verion 1.03 was then run with a distance threshold of 0.3 and samples grouped in to severe and non-severe groups.
```
corset -d 0.3 -g S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,S,I,I,I,I,I,I,I,I,I,I,I,I,I,I,I,I,I,I,I,I,I *severe.bam *non_severe.bam
```

##Load clustered transcript counts
We first load the resulting clusters and their respective counts into R.

```{r}
setwd(wd)

clusters <- read.delim("./data/clusters-0.3.txt", header=FALSE)
colnames(clusters) <- c("Transcript", "Cluster")

transcriptCounts <- read.table("./data/counts-0.3.txt", header=TRUE, sep="\t", row.names=1)
colnames(transcriptCounts) <- gsub("_S.*","",colnames(transcriptCounts))
```

##Remove low count samples
As was done previously we remove those samples with low coverage.

```{r}
setwd(wd)
transcriptCounts <- transcriptCounts[,!(colnames(transcriptCounts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
```

##PCA plot of Transcripts
Next we perfrom look at a PCA plot of the samples based on the log normalised counts of the reads assigned to each of the clusters. Clusters with low coverage, specifically less than 180 reads per million in at least 5 samples were filtered prior to performing PCA.

Notably unlike the all gene analysis the severe samples are more closely clustered together. In contrast the non-severe samples do not cluster so tightly. This is consistent with a more conserved set of genes being present in the severe malaria cases. A similar result was also observed in the combined assembly analysis.

```{r}
dge <- DGEList(counts=transcriptCounts)
cpm(10, mean(dge$samples$lib.size))
keep <- rowSums(cpm(dge)>180) >= 5
dge <- dge[keep,,keep.lib.sizes=FALSE]
categories <- as.factor(substring(rownames(dge$samples),1,1))
plotPCA(cpm(calcNormFactors(dge, method="RLE"), normalized.lib.sizes=TRUE, log=T), col=colors[categories], cex=0.8, isLog=TRUE)
```

##Differential Analysis
The default DESeq2 differential analysis pipeline is then used to investigate differential expression.

```{r}
#create design matrix
colData <- data.frame(disease=as.factor(substring(colnames(transcriptCounts),1,1)))
rownames(colData) <- colnames(transcriptCounts)

#create DESeq2 dataset
dds <- DESeqDataSetFromMatrix(countData = transcriptCounts
                            , colData = colData
                            , design = ~disease)

##estimate Sizefactors
dds <- estimateSizeFactors(dds)

#run DESeq2 pipeline
dds <- DESeq(dds)
res <- results(dds,contrast=c("disease","S","I"), pAdjustMethod = "BH")
summary(res, alpha=0.05)

Sig <- res[!is.na(res$padj),]
Sig <- Sig[Sig$padj<0.05,]

SigUp <- Sig[Sig$log2FoldChange>0,]
SigDown <- Sig[Sig$log2FoldChange<0,]
```

After Benjamini-Hochberg correction for multiple testing, `r dim(Sig)[1]` clusters were found to be differentially expressed. Of these `r dim(SigUp)[1]` were upregulated in severe disease.

##Heatmap of up-regulated transcripts
A heatmap of the log normalised samplewise read counts for each cluster is given below. It indicates that Cluster-10.1182 is expressed in the majority of samples. It also suggests that Clusters 10.583, 10.839 and 10.1147 are influenced by the same three severe samples.

```{r}
select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(SigUp),]

disease <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
colData$phenotype[colData$disease=="I"] <- "non-severe"
colData$phenotype[colData$disease=="S"] <- "severe"
colData$disease <- NULL

pheatmap(log2.norm.counts, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=TRUE, annotation_col=colData, annotation_colors = anno_colors)
```

##Heatmap of down-regulated transcripts
The larger number of clusters that are reported as downregulated may be a result of the increased diversity in the non-severe transcripts. The heatmap suggests there are a number of transcript clusters that are not present very often in the severe samples.

```{r}
select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(SigDown),]

pheatmap(log2.norm.counts, cluster_rows=TRUE, show_rownames=FALSE, cluster_cols=TRUE, annotation_col=colData, annotation_colors = anno_colors)
```

##Table of results
Finally a table of the results of the differential expression analysis is given for every cluster.

```{r}
#Collect sequence labels for each cluster
getTrans <- function(x) {
  return(paste(clusters$Transcript[clusters$Cluster==x], collapse=","))
}

final <- res
final$transcripts <- lapply(rownames(final), getTrans)
#Now print the table
final <- final[order(final$padj),]
final$padj <- format(final$padj, scientific=TRUE, digits=3)
final$pvalue <- format(final$pvalue, scientific=TRUE, digits=3)
kable(final, digits=3)
```

##Session Information
```{r}
sessionInfo()
```

### 2017-08-22 domain_level/TextS4_domain_tree_analysis.Rmd
renamed files
added in boxplot for domain and transcript expression

--- 2017-01-12 domain_level/domain_tree_analysis.Rmd
@@
 ##Get read counts for each domain
-Used featureCounts allowing for overlap. That is, if a read overlaps more than one segment it is counted as a hit for both segments.
+The assembled VAR DNA transcripts are first translated into protein sequence.
+This is done by translating each sequence into all six reading frames and keeping the frame with the lowest number of stop codons.
+If all of the six frames had more than 2 stop codons present the open reading frames of each of the six translations were taken and searched against the homology block models of Rask et al.
+An ORF was kept if it had at least two distinct homology blocks present.
+This approach accurately translated over 99% of the sequences.
+Upon manual inspection of the differentially expressed transcripts for the combined assembly, transcript 258591_soapGraphK61 was found to be translated in the reverse frame.
+This did not affect the domain analysis as the seperate assembly transcripts were used.
+
+Reads are aligned to the de novo assembled transcripts which are annotated according to the alignment of domain models with the respective protein transcripts.
+The VAR domains of Rask et al were aligned to the protein sequences using HMMER v3.1b1.
+
+```
+hmmsearch --nonull --noali --max -E 1e-8 --domtblout rask_domain_hmmerSearch.txt all_protein_exon1_domain_HMMs.hmm  ./data/seperate_assembly_transcipts_PROTEIN.fasta
+```
+
+An SAF file was created from the HMMER alignments and read counts were obtained using featureCounts allowing for overlap.
+That is, if a read overlaps more than one segment it is counted as a hit for both segments.
 ```{r}
@@
 ##A PCA plot of normalised counts for domains of Rask et al
-Reads are annotated to the de novo assembled transcripts which are annotated with the hidden Markov model domains profiles of Rask et al.
 ```{r}
@@
 pheatmap(log2.norm.counts, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=TRUE, annotation_col=annotData, annotation_colors = anno_colors)
+```
+
+A similar plot but in a different format
+```{r}
+norm_dom_counts <- counts(dds_dom, normalized=TRUE)
+sigUp_norm_counts <- norm_dom_counts[rownames(norm_dom_counts) %in% rownames(resSigLFC),]
+
+percent_counts <- t(t(sigUp_norm_counts)/colSums(norm_dom_counts))
+
+sigUp_norm_counts_df <- melt(percent_counts)
+colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of expression")
+sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
+sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
+colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of VAR expression", "Phenotype")
+
+gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=`Proportion of VAR expression`
+                                       , colour=Phenotype))
+gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
+gg <- gg + facet_wrap(~`Domain Cluster`) + theme_bw()
+gg <- gg + scale_colour_manual(values = colors)
+gg
 ```
@@
 
-From those nodes that are found to be significant after multiple testing correction we keep only the most signifcant from one path through the tree.
+From those nodes that are found to be significant after multiple testing correction we keep only the most significant from one path through the tree.
 That is if a node has an ancestor or child that has a smaller p-value, it will not be considered.
-This is achieved by iteratively selecting the most significant node and removing it's ancestors and children from the list of signifcant nodes.
+This is achieved by iteratively selecting the most significant node and removing it's ancestors and children from the list of significant nodes.
 
@@
 
+The boxplot indicating the same information in a different format.
+
+```{r}
+norm_counts <- counts(dds, normalized=TRUE)
+sigUp_norm_counts <- norm_counts[rownames(norm_counts) %in% rownames(resUp),]
+
+percent_counts <- t(t(sigUp_norm_counts)/colSums(norm_dom_counts))
+
+sigUp_norm_counts_df <- melt(percent_counts)
+colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of expression")
+sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
+sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
+colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of VAR expression", "Phenotype")
+
+gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=`Proportion of VAR expression`
+                                       , colour=Phenotype))
+gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
+gg <- gg + facet_wrap(~`Domain Cluster`) + theme_bw()
+gg <- gg + scale_colour_manual(values = colors)
+gg
+```
+
 ##Heatmap of down-regulated clusters
@@
 # }
-# plot_tree("670_X0.6", edges, p.vals, return_script=F)
+# plot_tree("780_X0.6", edges, p.vals, return_script=F)
 # writeTree("670_X0.6", "/home/users/allstaff/tonkin-hill.g/Dropbox/Var transcriptome manuscript/Figures/VAR differential analysis (4)/domainTrees/", edges, p.vals)
@@
 ```
+
+##Session Information
+```{r}
+sessionInfo()
+```

### 2017-08-22 transcript_level/TextS2_combined_assembly_transcript_analysis.Rmd
added in boxplot for domain and transcript expression

@@
 
+A similar plot but in a different format
+```{r}
+tran_counts <- counts(dds, normalized=TRUE)
+sigUp_norm_counts <- tran_counts[rownames(tran_counts) %in% rownames(resUp),]
+
+percent_counts <- t(t(sigUp_norm_counts)/colSums(tran_counts))
+
+sigUp_norm_counts_df <- melt(percent_counts)
+colnames(sigUp_norm_counts_df) <- c("Transcript", "Isolate", "Proportion of expression")
+sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
+sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
+colnames(sigUp_norm_counts_df) <- c("Transcript", "Isolate", "Proportion of VAR expression", "Phenotype")
+
+gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=`Proportion of VAR expression`
+                                       , colour=Phenotype))
+gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
+gg <- gg + facet_wrap(~Transcript, scales = "free_y") + theme_bw()
+gg <- gg + scale_colour_manual(values = colors)
+gg
+```
+
 ##Heatmap of down-regulated segment clusters

### 2017-08-22 transcript_level/TextS3_seperate_assembly_CORSET_analysis.Rmd
added in boxplot for domain and transcript expression

@@
 
+A similar plot but in a different format
+```{r}
+tran_counts <- counts(dds, normalized=TRUE)
+sigUp_norm_counts <- tran_counts[rownames(tran_counts) %in% rownames(SigUp),]
+
+percent_counts <- t(t(sigUp_norm_counts)/colSums(tran_counts))
+
+sigUp_norm_counts_df <- melt(percent_counts)
+colnames(sigUp_norm_counts_df) <- c("Transcript Clusters", "Isolate", "Proportion of expression")
+sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
+sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
+colnames(sigUp_norm_counts_df) <- c("Transcript Clusters", "Isolate", "Proportion of VAR expression", "Phenotype")
+
+gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=`Proportion of VAR expression`
+                                       , colour=Phenotype))
+gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
+gg <- gg + facet_wrap(~`Transcript Clusters`, scales = "free_y") + theme_bw()
+gg <- gg + scale_colour_manual(values = colors)
+gg
+```
+
 ##Heatmap of down-regulated transcripts

### 2017-08-22 domain_level/TextS4_domain_tree_analysis.Rmd
added in boxplot for domain and transcript expression

@@
 gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
-gg <- gg + facet_wrap(~`Domain Cluster`) + theme_bw()
+gg <- gg + facet_wrap(~`Domain Cluster`, scales = "free_y") + theme_bw()
 gg <- gg + scale_colour_manual(values = colors)

### 2017-08-24 domain_level/TextS4_domain_tree_analysis.Rmd
added in domain and trasnncript length data

@@
 ```{r}
-norm_dom_counts <- counts(dds_dom, normalized=TRUE)
-sigUp_norm_counts <- norm_dom_counts[rownames(norm_dom_counts) %in% rownames(resSigLFC),]
+domain_length_df <- fread("./data/domain_lengths.csv", data.table = FALSE)
+domain_length_matrix <- matrix(domain_length_df[,2], ncol=1,
+                               dimnames = list(row=domain_length_df[,1], col="len"))
+rpkm_dom <- counts(dds_dom, normalized=FALSE)
+dom_length_vector <- domain_length_matrix[rownames(rpkm_dom),]
+domain_lib_size <- colSums(rpkm_dom)
+rpkm_dom <- (t(t(rpkm_dom)/domain_lib_size)/dom_length_vector)*10^9
+rpkm_dom <-  rpkm_dom[rownames(rpkm_dom) %in% rownames(resSigLFC),]
 
-percent_counts <- t(t(sigUp_norm_counts)/colSums(norm_dom_counts))
+# norm_dom_counts <- counts(dds_dom, normalized=TRUE)
+# sigUp_norm_counts <- norm_dom_counts[rownames(norm_dom_counts) %in% rownames(resSigLFC),]
+# 
+# percent_counts <- t(t(sigUp_norm_counts)/colSums(norm_dom_counts))
 
-sigUp_norm_counts_df <- melt(percent_counts)
-colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of expression")
+rpkm_dom <- rpkm_dom+1 #to deal with log transformation
+sigUp_norm_counts_df <- melt(rpkm_dom)
+colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "RPKM")
 sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
 sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
-colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of VAR expression", "Phenotype")
+colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "RPKM", "Phenotype")
 
-gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=`Proportion of VAR expression`
+gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=`RPKM`
                                        , colour=Phenotype))
-gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
-gg <- gg + facet_wrap(~`Domain Cluster`, scales = "free_y") + theme_bw()
-gg <- gg + scale_colour_manual(values = colors)
+gg <- gg + geom_boxplot(outlier.shape = NA, position = "dodge") + geom_jitter(width = 0.2, height = 0)
+gg <- gg + facet_wrap(~`Domain Cluster`, nrow=1)
+gg <- gg + scale_colour_manual(values = colors) + scale_y_log10()
+gg <- gg + theme_bw()
 gg

### 2017-08-24 domain_level/TextS4_domain_tree_analysis.Rmd
nearly done new domain graph

@@
 ```{r}
-norm_counts <- counts(dds, normalized=TRUE)
-sigUp_norm_counts <- norm_counts[rownames(norm_counts) %in% rownames(resUp),]
+rpkm_dom_clus <- counts(dds, normalized=FALSE)
+rpkm_dom_clus <- rpkm_dom_clus[rownames(rpkm_dom_clus) %in% rownames(resUp),]
+dom_length_vector <- domain_length_matrix[cluster_domain_type[rownames(rpkm_dom_clus)],]
+rownames(rpkm_dom_clus) <- paste(rownames(rpkm_dom_clus),
+                           cluster_domain_type[rownames(rpkm_dom_clus)] ,
+                           sep=" - ")
+rpkm_dom_clus <- (t(t(rpkm_dom_clus)/domain_lib_size)/dom_length_vector)*10^9
 
-percent_counts <- t(t(sigUp_norm_counts)/colSums(norm_dom_counts))
+rpkm_dom_clus <- rpkm_dom_clus+1 #to deal with log transformation
+sigUp_norm_counts_clus_df <- melt(rpkm_dom_clus)
+colnames(sigUp_norm_counts_clus_df) <- c("Domain Cluster", "Isolate", "RPKM")
 
-sigUp_norm_counts_df <- melt(percent_counts)
-colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of expression")
-sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
-sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
-colnames(sigUp_norm_counts_df) <- c("Domain Cluster", "Isolate", "Proportion of VAR expression", "Phenotype")
+sigUp_norm_counts_clus_df$disease[grepl("^S.*",sigUp_norm_counts_clus_df$Isolate)]<-"severe"
+sigUp_norm_counts_clus_df$disease[!grepl("^S.*",sigUp_norm_counts_clus_df$Isolate)]<-"non-severe"
+colnames(sigUp_norm_counts_clus_df) <- c("Domain Cluster", "Isolate", "RPKM", "Phenotype")
 
-gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=`Proportion of VAR expression`
+gg <- ggplot(sigUp_norm_counts_clus_df, aes(x=Phenotype, y=RPKM
                                        , colour=Phenotype))
 gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
-gg <- gg + facet_wrap(~`Domain Cluster`, scales = "free_y") + theme_bw()
-gg <- gg + scale_colour_manual(values = colors)
+gg <- gg + facet_wrap(~`Domain Cluster`) + theme_bw()
+gg <- gg + scale_colour_manual(values = colors) + scale_y_log10()
 gg

### 2017-11-08 domain_level/HmmerUproc.Rmd
added in hmmer uproc analysis in response to reviewer comments

---
title: "HMMER & UPROC domain analysis"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

###Load Libraries
```{r}
library(ggplot2)
library(reshape2)
library(edgeR)
library(plyr)
library(scales)

library(Rsubread)
library(structSSI)
library(data.table)
library(igraph)

library(cluster)
library(fastcluster)
library(RFLPtools)

library(pheatmap)
library(EDASeq)
library(dynamicTreeCut)
library(moduleColor)
library(knitr)

library(DESeq2)

wd = "./"
colors <- c("#0571b0","#ca0020")
```

Here we make use of UPROC to provide an intial filter of the reads before making user of HMMER to allocate the reads more accurately to domain models. The code used to run this pipeline is available from https://github.com/PapenfussLab/reads_to_domains.

```{bash, eval=FALSE}
parallel -j 5 /home/users/allstaff/tonkin-hill.g/transcriptome_manuscript_response/hmmerUproc/run_UPROCHMMERR.sh ::: /stornext/Bioinf/data/bioinf-data/Papenfuss_lab/projects/malaria/duffy_lab/clinical_isolate_RNA_seq_severe_nonsevere/data/nonsevere/I*_R1_*.fastq

parallel -j 5 /home/users/allstaff/tonkin-hill.g/transcriptome_manuscript_response/hmmerUproc/run_UPROCHMMERR.sh ::: /stornext/Bioinf/data/bioinf-data/Papenfuss_lab/projects/malaria/duffy_lab/clinical_isolate_RNA_seq_severe_nonsevere/data/severe/S*_R1_*.fastq
```

The script run_UPROCHMMERR.sh is copied for reference below
```{bash, eval=FALSE}
set -e

read1=$1

read2=${read1/_R1_/_R2_}

echo "running allocate_reads.py ${read1} ${read2}"

python ~/get_domains/reads_to_domains/reads_to_domains/allocate_reads.py -r $read1 -R $read2 -E 1e-2 -o /home/users/allstaff/tonkin-hill.g/transcriptome_manuscript_response/hmmerUproc/
```

##Process resulting counts

First load data into a count matrix
```{r}
files <- Sys.glob("./data/hmmerUproc/*.csv")
domain_count_list <- lapply(files, fread, data.table=FALSE, header=FALSE)
domain_names <- unique(unlist(lapply(domain_count_list, unlist(function(x) x[,1]))))
sample_names <- gsub("_S.*", "", gsub(".*Uproc/", "", files))
unique_sample_names <- unique(sample_names)
domain_counts <- matrix(0, nrow=length(domain_names), ncol=length(unique_sample_names),
                        dimnames = list(domain_names,unique_sample_names))
for (i in 1:length(domain_count_list)){
  domain_counts[domain_count_list[[i]][,1],sample_names[[i]]] <- domain_count_list[[i]][,2]
}

```

Now we can process it using the default DESeq2 pipeline as was done for the aligned Rask domain counts.

Again we remove samples SFD1, SFM9 and SFC.025 as they were not sequenced to a sufficient depth.
```{r}
setwd(wd)
domain_counts <- domain_counts[,!(colnames(domain_counts) %in% c('SFD1-CM', 'SFM9', 'SFC-025'))]
```

```{r}
x_domains <- DGEList(counts=domain_counts)
x_domains <- calcNormFactors(x_domains, method="RLE")
categories <- as.factor(substring(rownames(x_domains$samples),1,1))
```

###A PCA plot of normalised counts
```{r}
plotPCA(cpm(x_domains,normalized.lib.sizes=TRUE, log=T, prior.count=0.5), col=colors[categories], cex=0.8, isLog=TRUE)
title("VAR PCA - Previous Domains")
```

###Investigate differential expression
```{r}
colData <- data.frame(disease=categories)
rownames(colData) <- colnames(x_domains$counts)
dds_dom <- DESeqDataSetFromMatrix(countData = x_domains$counts
                              , colData = colData
                              , design = ~disease)
dds_dom <- estimateSizeFactors(dds_dom)
dds_dom <- DESeq(dds_dom) 
res <- results(dds_dom,contrast=c("disease","S","I"), pAdjustMethod = "BH")
summary(res, alpha=0.05)
```

##A heatmap of the Up-regulated domains
```{r}
res_domains <- res
resOrdered <- res_domains[order(res_domains$padj),]
resSig <- subset(resOrdered, padj < 0.05)

resSigLFC <- data.frame(subset(resSig, log2FoldChange > 0))

select <- order(rowMeans(counts(dds_dom, normalized=TRUE)), decreasing=TRUE)

nt <- normTransform(dds_dom)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resSigLFC),]
annotData <- colData

disease        <- colors
names(disease) <- c("non-severe", "severe")
anno_colors <- list(phenotype = disease)
annotData$phenotype[annotData$disease=="I"] <- "non-severe" 
annotData$phenotype[annotData$disease=="S"] <- "severe"
# annotData <- annotData[annotData$disease=="S",]
annotData$disease <- NULL
pheatmap(log2.norm.counts, cluster_rows=TRUE, show_rownames=TRUE, cluster_cols=TRUE, annotation_col=annotData, annotation_colors = anno_colors)
```

### 2017-11-08 domain_level/TextS4_domain_tree_analysis.Rmd
added in hmmer uproc analysis in response to reviewer comments

@@
 
-gg <- ggplot(sigUp_norm_counts_clus_df, aes(x=Phenotype, y=RPKM
+combined_rask_new_df <- rbind(sigUp_norm_counts_clus_df, sigUp_norm_counts_df)
+
+gg <- ggplot(combined_rask_new_df, aes(x=Phenotype, y=RPKM
                                        , colour=Phenotype))
-gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.2, height = 0)
-gg <- gg + facet_wrap(~`Domain Cluster`) + theme_bw()
+gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.05, height = 0)
+gg <- gg + facet_wrap(~`Domain Cluster`, nrow=1)
 gg <- gg + scale_colour_manual(values = colors) + scale_y_log10()
+gg <- gg + theme_bw()
+gg <- gg +  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
+gg <- gg + theme(panel.spacing.x=unit(0, "lines"),panel.spacing.y=unit(1, "lines"))
+gg <- gg + theme(strip.text.x = element_text(angle = 90))
 gg

### 2017-11-08 transcript_level/TextS2_combined_assembly_transcript_analysis.Rmd
added in hmmer uproc analysis in response to reviewer comments

--- 2017-03-04 transcript_level/TextS2_combined_assembly_transcript_analysis.Rmd
@@
 
+A similar plot but in a different format
+```{r}
+combined_length_df <- fread("./data/combined_assembly_transcript_lengths.csv"
+                            , data.table = FALSE)
+combined_length_matrix <- matrix(combined_length_df[,2], ncol=1,
+                               dimnames = list(row=combined_length_df[,1],
+                                               col="len"))
+rpkm_combined <- counts(dds, normalized=FALSE)
+combined_length_vector <- combined_length_matrix[rownames(rpkm_combined),]
+combined_lib_size <- colSums(rpkm_combined)
+rpkm_combined <- (t(t(rpkm_combined)/combined_lib_size)/combined_length_vector)*10^9
+rpkm_combined <-  rpkm_combined[rownames(rpkm_combined) %in% rownames(resUp),]
+rpkm_combined <- rpkm_combined + 1
+# combined_rask_new_df <- counts(dds, normalized=TRUE)
+# sigUp_norm_counts <- tran_counts[rownames(tran_counts) %in% rownames(resUp),]
+# 
+# percent_counts <- t(t(sigUp_norm_counts)/colSums(tran_counts))
+
+sigUp_norm_counts_df <- melt(rpkm_combined)
+colnames(sigUp_norm_counts_df) <- c("Transcript", "Isolate", "RPKM")
+sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
+sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
+colnames(sigUp_norm_counts_df) <- c("Transcript", "Isolate", "RPKM", "Phenotype")
+
+gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=RPKM
+                                       , colour=Phenotype))
+gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.05, height = 0)
+gg <- gg + facet_wrap(~Transcript, nrow=1)
+gg <- gg + scale_colour_manual(values = colors) + scale_y_log10()
+gg <- gg + theme_bw()
+gg <- gg +  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
+gg <- gg + theme(panel.spacing.x=unit(0, "lines"),panel.spacing.y=unit(1, "lines"))
+gg <- gg + theme(strip.text.x = element_text(angle = 90))
+gg
+```
+
 ##Heatmap of down-regulated segment clusters

### 2017-11-08 transcript_level/TextS3_seperate_assembly_CORSET_analysis.Rmd
added in hmmer uproc analysis in response to reviewer comments

--- 2017-03-04 transcript_level/TextS3_seperate_assembly_CORSET_analysis.Rmd
@@
 
+A similar plot but in a different format
+```{r}
+clustered_length_df <- fread("./data/sig_clusters_combined_length.csv"
+                            , data.table = FALSE)
+clustered_length_matrix <- matrix(clustered_length_df[,2], ncol=1,
+                               dimnames = list(row=clustered_length_df[,1],
+                                               col="len"))
+rpkm_clustered <- counts(dds, normalized=FALSE)
+clustered_lib_size <- colSums(rpkm_clustered)
+
+rpkm_clustered <- rpkm_clustered[rownames(rpkm_clustered) %in% rownames(SigUp),]
+
+clustered_length_vector <- clustered_length_matrix[rownames(rpkm_clustered),]
+
+rpkm_clustered <- (t(t(rpkm_clustered)/clustered_lib_size)/clustered_length_vector)*10^9
+rpkm_clustered <- rpkm_clustered + 1
+
+# tran_counts <- counts(dds, normalized=TRUE)
+# sigUp_norm_counts <- tran_counts[rownames(tran_counts) %in% rownames(SigUp),]
+# 
+# percent_counts <- t(t(sigUp_norm_counts)/colSums(tran_counts))
+
+sigUp_norm_counts_df <- melt(rpkm_clustered)
+colnames(sigUp_norm_counts_df) <- c("Transcript", "Isolate", "RPKM")
+sigUp_norm_counts_df$disease[grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"severe"
+sigUp_norm_counts_df$disease[!grepl("^S.*",sigUp_norm_counts_df$Isolate)]<-"non-severe"
+colnames(sigUp_norm_counts_df) <- c("Transcript Clusters", "Isolate", "RPKM", "Phenotype")
+
+gg <- ggplot(sigUp_norm_counts_df, aes(x=Phenotype, y=RPKM
+                                       , colour=Phenotype))
+gg <- gg + geom_boxplot(outlier.shape = NA) + geom_jitter(width = 0.05, height = 0)
+gg <- gg + facet_wrap(~`Transcript Clusters`, nrow=1)
+gg <- gg + scale_colour_manual(values = colors) + scale_y_log10()
+gg <- gg + theme_bw()
+gg <- gg +  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
+gg <- gg + theme(panel.spacing.x=unit(0, "lines"),panel.spacing.y=unit(1, "lines"))
+gg <- gg + theme(strip.text.x = element_text(angle = 90))
+gg
+```
+
 ##Heatmap of down-regulated transcripts

### 2017-11-08 all_gene_analysis/all_gene_analysis_additional_treatment_removed.Rmd
added in analysis with additional isolates excluded that had previously been treated. This was in response to reviewer comments

--- 2017-03-04 all_gene_analysis/TextS1_all_gene_analysis.Rmd
@@
 ---
-title: "All Gene Supplementary Material"
+title: "All Gene Supplementary Material - Additional samples removed."
 author: "Gerry Tonkin-Hill"
@@
 Samples SFU2 and SFU3 were drug treated and are removed.
+We repeated the analysis of the all gene DGE omitting a few patients who had anti-malarial treatments in preceding months.
+The patients omitted are: SFM 007, SFC 023, IFM 012 and IFM 021.
 ```{r}
 x$counts <- x$counts[,!grepl("SFU2|SFU\\.3", colnames(x$counts))]
+x$counts <- x$counts[,!grepl("SFM\\.7|SFC\\.023|IFM12|IFM21", colnames(x$counts))]
 filt.annnotation <- fc2$annotation[fc2$annotation$GeneID %in% rownames(x$counts),]
 dge <- DGEList(counts=x$counts, genes=filt.annnotation)
-
 ```

### 2017-11-08 prediction_analysis_for_reviewer/var_discrimination_review_response.Rmd
added in prediction analysis as requested from reviewer

---
title: "Determining descrimination of VAR features"
author: "Gerry Tonkin-Hill"
date: "`r Sys.Date()`"
output: 
  html_document:
    fig_width: 12
    fig_height: 8
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

###Load Libraries
```{r}
#libraries used for analysis
library(edgeR)
library(limma)
library(quadprog)
library(ruv)

#libraries required for graphics
library(ggplot2)
library(reshape2)
library(EDASeq)
library(RColorBrewer)
library(VennDiagram)
library(data.table)
library(dplyr)
library(stringr)
library(knitr)
library(pheatmap)
library(glmnet)

wd = "./"
colors <- c("#0571b0","#ca0020")
```

To investigate the utility of using the identified features to distinguish severe and non-severe cases we took the up-regulated features in each category (transcript, domain or segment) and performed logistic regression with lasso regularization. We used cross-validation to determine the optimal $\lamda$ value and give an indication of the best case discrimination between the cases. It should be noted that the classification accuracy reported here may not generalise to other samples. Due to the lack of additional samples a more extensive validation was not possible.

##Transcript level

###Combined Assembly Transcripts

```{r}
# knit('../transcript_level/TextS2_combined_assembly_transcript_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS2_combined_assembly_transcript_analysis.R')
setwd("../transcript_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS2_combined_assembly_transcript_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]
```

###Transcript Clusters
```{r}
rm(list=ls(all=TRUE))

# knit('../transcript_level/TextS3_seperate_assembly_CORSET_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS3_seperate_assembly_CORSET_analysis.R')
setwd("../transcript_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS3_seperate_assembly_CORSET_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(SigUp),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]
```

##Domain Level
```{r}
rm(list=ls(all=TRUE))

# knit('../domain_level/TextS4_domain_tree_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS4_domain_tree_analysis.R')
setwd("../domain_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS4_domain_tree_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds_dom, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds_dom)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resSigLFC),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]

x2 <- t(log2.norm.counts)
y2 <- substring(colnames(log2.norm.counts),1,1)
y2[y2=="I"] <- "non-severe"
y2[y2=="S"] <- "severe"
y2 <- as.factor(y2)
cvfit2 = cv.glmnet(x2, y2, family = "binomial", type.measure = "class")
plot(cvfit2)
cvfit2$lambda.1se
coef(cvfit2, s = "lambda.1se")
cvfit2$lambda.min
coef(cvfit2, s = "lambda.min")
cvfit2$cvm[cvfit2$lambda==cvfit2$lambda.min]
```

##Segment Level

```{r}
rm(list=ls(all=TRUE))

# knit('../segment_level/TextS6_segment_level_analysis.Rmd',
#      tangle=TRUE,
#      output = './TextS6_segment_level_analysis.R')
setwd("../segment_level/")
pdf(file = NULL)
source('../prediction_analysis_for_reviewer/TextS6_segment_level_analysis.R')
dev.off()
setwd("../prediction_analysis_for_reviewer/")

select <- order(rowMeans(counts(dds_block, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds_block)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resSigLFC),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]

select <- order(rowMeans(counts(dds, normalized=TRUE)), decreasing=TRUE)
nt <- normTransform(dds)
log2.norm.counts <- assay(nt)[select,]
log2.norm.counts <- assay(nt)[rownames(nt) %in% rownames(resUp),]

x <- t(log2.norm.counts)
y <- substring(colnames(log2.norm.counts),1,1)
y[y=="I"] <- "non-severe"
y[y=="S"] <- "severe"
y <- as.factor(y)
cvfit = cv.glmnet(x, y, family = "binomial", type.measure = "class")
plot(cvfit)
cvfit$lambda.1se
coef(cvfit, s = "lambda.1se")
cvfit$lambda.min
coef(cvfit, s = "lambda.min")
cvfit$cvm[cvfit$lambda==cvfit$lambda.min]
```

### 2017-11-08 prediction_analysis_for_reviewer/TextS2_combined_assembly_transcript_analysis.R

@@ -1,0 +1,210 @@
+library(limma)
+library(edgeR)
+library(DESeq2)
+library(genefilter)
+library(EDASeq)
+library(RColorBrewer)
+library(VennDiagram)
+library(knitr)
+library(reshape2)
+
+library(pheatmap)
+library(dynamicTreeCut)
+library(moduleColor)
+library(ggplot2)
+
+wd = "./"
+colors <- c("#0571b0","#ca0020")
+
+## ------------------------------------------------------------------------
+setwd(wd)
+load("./data/geneWithoutVivax_and_VARgene_countDB_subread_uH.RData")
+var_gene_names <- read.delim("./data/var_gene_names.txt")
+var_gene_names <- var_gene_names$X.Gene.ID.
+#Load Non-var_genes
+fc2 <- fc_genes_aligned_w_Pfhuman_withoutVivax
+t <- fc2$counts
+
+#Rename samples to something more sensible
+# colnames(t) <- gsub("alignment\\.sam","",colnames(t))
+# colnames(t) <- gsub(".*vivax\\.","",colnames(t))
+# colnames(t) <- gsub("_S.*","", colnames(t))
+nonVar <- t
+nonVar.annot <- fc2$annotation
+
+var <- var_counts$counts
+var.annot <- var_counts$annotation
+# colnames(var) <- gsub(".*alignments\\.","",colnames(var))
+# colnames(var) <- gsub("alignment.sam","",colnames(var))
+# colnames(var) <- gsub("_L00.*", "", colnames(var))
+# colnames(var) <- gsub("_S.*", "", colnames(var))
+var_names <- rownames(var)
+
+stopifnot(colnames(var)==colnames(nonVar))
+combined_counts <- rbind(nonVar, var)
+combined_annot <- rbind(nonVar.annot, var.annot)
+
+## ------------------------------------------------------------------------
+combined_counts <- combined_counts[,!(colnames(combined_counts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
+
+## ------------------------------------------------------------------------
+stopifnot(rownames(combined_counts)==combined_annot$GeneID)
+dge <- DGEList(counts=combined_counts, genes=combined_annot)
+dge_combined <- dge
+
+cpm(10, mean(dge_combined$samples$lib.size))
+
+keep <- rowSums(cpm(dge)>1) >= 5
+dge <- dge[keep,,keep.lib.sizes=FALSE]
+
+categories <- as.factor(substring(colnames(combined_counts),1,1))
+
[…]

### 2017-11-08 prediction_analysis_for_reviewer/TextS3_seperate_assembly_CORSET_analysis.R

@@ -1,0 +1,138 @@
+
+library(DESeq2)
+library(ggplot2)
+library(edgeR)
+library(knitr)
+library(pheatmap)
+library(EDASeq)
+
+wd <- "./"
+colors <- c("#0571b0","#ca0020")
+
+## ------------------------------------------------------------------------
+setwd(wd)
+
+clusters <- read.delim("./data/clusters-0.3.txt", header=FALSE)
+colnames(clusters) <- c("Transcript", "Cluster")
+
+transcriptCounts <- read.table("./data/counts-0.3.txt", header=TRUE, sep="\t", row.names=1)
+colnames(transcriptCounts) <- gsub("_S.*","",colnames(transcriptCounts))
+
+## ------------------------------------------------------------------------
+setwd(wd)
+transcriptCounts <- transcriptCounts[,!(colnames(transcriptCounts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
+
+## ------------------------------------------------------------------------
+dge <- DGEList(counts=transcriptCounts)
+cpm(10, mean(dge$samples$lib.size))
+keep <- rowSums(cpm(dge)>180) >= 5
+dge <- dge[keep,,keep.lib.sizes=FALSE]
+categories <- as.factor(substring(rownames(dge$samples),1,1))
+plotPCA(cpm(calcNormFactors(dge, method="RLE"), normalized.lib.sizes=TRUE, log=T), col=colors[categories], cex=0.8, isLog=TRUE)
+
+## ------------------------------------------------------------------------
+#create design matrix
+colData <- data.frame(disease=as.factor(substring(colnames(transcriptCounts),1,1)))
+rownames(colData) <- colnames(transcriptCounts)
+
+#create DESeq2 dataset
+dds <- DESeqDataSetFromMatrix(countData = transcriptCounts
+                            , colData = colData
+                            , design = ~disease)
+
+##estimate Sizefactors
+dds <- estimateSizeFactors(dds)
+
+#run DESeq2 pipeline
+dds <- DESeq(dds)
+res <- results(dds,contrast=c("disease","S","I"), pAdjustMethod = "BH")
+summary(res, alpha=0.05)
+
+Sig <- res[!is.na(res$padj),]
+Sig <- Sig[Sig$padj<0.05,]
+
+SigUp <- Sig[Sig$log2FoldChange>0,]
+SigDown <- Sig[Sig$log2FoldChange<0,]
+
[…]

### 2017-11-08 prediction_analysis_for_reviewer/TextS4_domain_tree_analysis.R

@@ -1,0 +1,595 @@
+
+library(ggplot2)
+library(reshape2)
+library(edgeR)
+library(plyr)
+library(scales)
+
+library(Rsubread)
+library(structSSI)
+library(data.table)
+library(igraph)
+
+library(cluster)
+library(fastcluster)
+library(RFLPtools)
+
+library(pheatmap)
+library(EDASeq)
+library(dynamicTreeCut)
+library(moduleColor)
+library(knitr)
+
+library(DESeq2)
+
+wd = "./"
+colors <- c("#0571b0","#ca0020")
+
+## ------------------------------------------------------------------------
+setwd(wd)
+# files <- Sys.glob("/home/users/allstaff/tonkin-hill.g/gene_level_var_analysis/alignments_seperate_assemblies_best1/*.sam")
+# dom_counts <- featureCounts(files, annot.ext = "./data/NTS_DBL_CIDR_seperate_assembly_domains.SAF"
+#                             , isPairedEnd=TRUE, isGTFAnnotationFile=FALSE
+#                             , countChimericFragments=TRUE
+#                             , allowMultiOverlap=TRUE
+#                             , useMetaFeatures=FALSE
+#                             , nthreads=20
+#                             , countMultiMappingReads=TRUE)
+# 
+# #combine counts for different runs of the same sample
+# colnames(dom_counts$counts) <- gsub(".*_best1\\.","",colnames(dom_counts$counts))
+# colnames(dom_counts$counts) <- gsub("_S.*","",colnames(dom_counts$counts))
+# temp <- dom_counts
+# temp$counts <- t(rowsum(t(temp$counts), group = rownames(t(temp$counts))))
+# dim(temp$counts)
+# dom_counts <- temp
+# # #
+# saveRDS(dom_counts, "./data/dom_counts.rds")
+
+dom_counts <- readRDS("./data/dom_counts.rds")
+
+## ------------------------------------------------------------------------
+setwd(wd)
+dom_counts$counts <- dom_counts$counts[,!(colnames(dom_counts$counts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
+
+## ------------------------------------------------------------------------
+setwd(wd)
+prev_dom_counts <- data.frame(dom_counts$counts)
[…]

### 2017-11-08 prediction_analysis_for_reviewer/TextS6_segment_level_analysis.R

@@ -1,0 +1,493 @@
+
+library(DESeq2)
+library(ggplot2)
+library(edgeR)
+library(Rsubread)
+library(data.table)
+library(RFLPtools)
+library(EDASeq)
+library(igraph)
+library(knitr)
+library(pheatmap)
+library(plyr)
+library(dplyr)
+library(stringr)
+library(dynamicTreeCut)
+library(moduleColor)
+
+colors <- c("#0571b0","#ca0020")
+wd <- "./"
+
+## ------------------------------------------------------------------------
+setwd(wd)
+# files <- Sys.glob("/home/users/allstaff/tonkin-hill.g/gene_level_var_analysis/alignments_seperate_assemblies_best1/*.sam")
+# block_counts <- featureCounts(files, annot.ext = "./data/rask_blocks/rask_blocks_in_seperate_assembly.SAF"
+#                             , isPairedEnd=TRUE, isGTFAnnotationFile=FALSE
+#                             , countChimericFragments=TRUE
+#                             , allowMultiOverlap=TRUE
+#                             , useMetaFeatures=FALSE
+#                             , nthreads=20)
+# 
+# #combine counts for different runs of the same sample
+# colnames(block_counts$counts) <- gsub(".*_best1\\.","",colnames(block_counts$counts))
+# colnames(block_counts$counts) <- gsub("_S.*","",colnames(block_counts$counts))
+# temp <- block_counts
+# temp$counts <- t(rowsum(t(temp$counts), group = rownames(t(temp$counts))))
+# dim(temp$counts)
+# rask_block_counts <- temp
+# 
+# saveRDS(rask_block_counts, "./data/rask_block_counts.rds")
+
+rask_block_counts <- readRDS("./data/rask_block_counts.rds")
+
+## ------------------------------------------------------------------------
+rask_block_counts$counts <- rask_block_counts$counts[,!(colnames(rask_block_counts$counts) %in% c('SFD1.CM', 'SFM9', 'SFC.025'))]
+
+## ------------------------------------------------------------------------
+prev_block_counts <- data.frame(rask_block_counts$counts)
+prev_block_counts$blockType <- unlist(lapply(rownames(prev_block_counts), function(x){
+  temp <- unlist(strsplit(x, "_"))
+  return(paste("block",temp[length(temp)-1],sep="_"))
+  }))
+prev_block_counts <- ddply(prev_block_counts, "blockType", numcolwise(sum))
[…]

### 2017-12-23 all_gene_analysis/all_gene_analysis_additional_treatment_removed.Rmd
added in gene set tests

@@
 We repeated the analysis of the all gene DGE omitting a few patients who had anti-malarial treatments in preceding months.
-The patients omitted are: SFM 007, SFC 023, IFM 012 and IFM 021.
+The patients omitted are: SFC 023, SFM007, IFM 012 and IFM 021.
+
 ```{r}
 x$counts <- x$counts[,!grepl("SFU2|SFU\\.3", colnames(x$counts))]
-x$counts <- x$counts[,!grepl("SFM\\.7|SFC\\.023|IFM12|IFM21", colnames(x$counts))]
+x$counts <- x$counts[,!grepl("SFC\\.023|SFM\\.7|IFM12|IFM21", colnames(x$counts))]
 filt.annnotation <- fc2$annotation[fc2$annotation$GeneID %in% rownames(x$counts),]
@@
 
+##Gene set analysis
+
+First GO set enrichment using limma's mroast.
+
+```{r}
+# genesets <- AnnotationDbi::select(org.Pf.plasmo.db, keys=top_ring_ruv$GeneID, keytype="ALIAS", 
+#        columns = c("GO","PATH"))
+# go.gene.sets <- lapply(split.data.frame(genesets,genesets$GO), function(x) as.vector(x$ALIAS))
+# kegg.gene.sets <- lapply(split.data.frame(genesets,genesets$PATH), function(x) as.vector(x$ALIAS))
+MIN_SET_SIZE <- 3
+
+go.gene.df <- fread("./data/go_gene_lists.txt", sep="\t", data.table = FALSE)
+go.gene.sets <- lapply(go.gene.df$Genes, function(x) {
+  x <- unlist(str_split(x, ","))
+  x <- x[x!=""]
+  })
+names(go.gene.sets) <- go.gene.df$GO_ID
+go.gene.sets <- go.gene.sets[unlist(lapply(go.gene.sets, length))>=MIN_SET_SIZE]
+
+kegg.gene.df <- fread("./data/kegg_gene_lists.txt", sep="\t", data.table = FALSE)
+kegg.gene.sets <- lapply(kegg.gene.df$Genes, function(x) {
+  x <- unlist(str_split(x, ","))
+  x <- x[x!=""]
+  })
+names(kegg.gene.sets) <- kegg.gene.df$KEGG_ID
+kegg.gene.sets <- kegg.gene.sets[unlist(lapply(kegg.gene.sets, length))>=MIN_SET_SIZE]
+
+indices <- ids2indices(go.gene.sets, rownames(v2), remove.empty=TRUE)
+go_mr <- mroast(v2, index=indices, design=modRUV, contrast=2,
+                nrot=10000, set.statistic="floormean")
+go_cam <- camera(v2, index=indices, design=modRUV, contrast=2)
+kable(go_mr, digits=3)
+```
+
+Also test using limma's camera
+
+```{r}
+kable(go_cam, digits=3)
+```
+
+Now KEGG pathway enrichment again making use of mroast and camera.
+```{r}
+indices <- ids2indices(kegg.gene.sets, rownames(v2))
+keg_mr <- mroast(v2, index=indices, design=modRUV, contrast=2, nrot=10000,
+                 set.statistic="floormean")
+keg_cam <- camera(v2, index=indices, design=modRUV, contrast=2)
+kable(keg_mr, digits=3)
+```
+
+Also test using limma's camera
+
+```{r}
+kable(keg_cam, digits=3)
+```
+
 ##Gene sets from Daily et al and Shock et al
@@
 
-##KEGG analysis
-```{r}
-# setwd(wd)
-# 
-# entrez_de <- entrez_geneID_mappings$`[Entrez Gene ID]`[entrez_geneID_mappings$`[Gene ID]` %in% top_ring_ruv$GeneID]
-# entrez_de <- entrez_de[entrez_de!="null"]
-# entrez_de <- str_split_fixed(entrez_de, pattern = ",", n = 2)[,1]
-# 
-# kegg <- kegga(entrez_de, species.KEGG="pfa", convert=TRUE)
-# kable(topKEGG(kegg, n=Inf))
-```
-
-##GO analysis
-```{r}
-# go <- goana(entrez_de, species.KEGG="pfa", convert=TRUE)
-# kable(topGO(go, n=100))
-```
-
 ####System Information

### 2017-12-23 all_gene_analysis/all_gene_analysis_additional_treatment_removed.Rmd
updated changes to new supp file

@@
 go_cam <- camera(v2, index=indices, design=modRUV, contrast=2)
+
+go_mr$PValue <- format(go_mr$PValue, scientific=TRUE, digits=3)
+go_mr$FDR <- format(go_mr$FDR, scientific=TRUE, digits=3)
+go_mr$PValue.Mixed <- format(go_mr$PValue.Mixed, scientific=TRUE, digits=3)
+go_mr$FDR.Mixed <- format(go_mr$FDR.Mixed, scientific=TRUE, digits=3)
 kable(go_mr, digits=3)
@@
 ```{r}
-kable(go_cam, digits=3)
+go_cam$PValue <- format(go_cam$PValue, scientific=TRUE, digits=3)
+go_cam$FDR <- format(go_cam$FDR, scientific=TRUE, digits=3)
+kable(go_cam)
 ```
@@
 keg_cam <- camera(v2, index=indices, design=modRUV, contrast=2)
+
+keg_mr$PValue <- format(keg_mr$PValue, scientific=TRUE, digits=3)
+keg_mr$FDR <- format(keg_mr$FDR, scientific=TRUE, digits=3)
+keg_mr$PValue.Mixed <- format(keg_mr$PValue.Mixed, scientific=TRUE, digits=3)
+keg_mr$FDR.Mixed <- format(keg_mr$FDR.Mixed, scientific=TRUE, digits=3)
 kable(keg_mr, digits=3)
@@
 ```{r}
+keg_cam$PValue <- format(keg_cam$PValue, scientific=TRUE, digits=3)
+keg_cam$FDR <- format(keg_cam$FDR, scientific=TRUE, digits=3)
 kable(keg_cam, digits=3)

### 2018-01-28 all_gene_analysis/all_gene_analysis_additional_treatment_removed.Rmd
added in aditional scripts

--- 2017-11-08 all_gene_analysis/all_gene_analysis_additional_treatment_removed.Rmd
@@
 
+##Gene set analysis
+
+First we can set up the gene sets after downloading the set annotations from PlasmoDB.
+```{r}
+MIN_SET_SIZE <- 3
+
+#GO sets
+go.data <- fread("./data/PlasmoDB-35_Pfalciparum3D7_GO.gaf", skip = 1, sep="\t",
+                         data.table = FALSE)
+go.data <- go.data[go.data$V9=="P",] #Looking at biological processes
+go.terms <- unique(go.data$V5)
+go.gene.sets <- lapply(go.terms, function(go){
+  unique(go.data$V2[grepl(go, go.data$V5, fixed = TRUE)])
+})
+names(go.gene.sets) <- go.terms
+go.gene.sets <- go.gene.sets[unlist(lapply(go.gene.sets, length))>=MIN_SET_SIZE]
+
+#KEGG sets
+kegg.annot.data <- fread("./data/KEGG_Metacyc_PlasmoDB.csv", 
+                         data.table = FALSE)
+kegg.annot.data <- kegg.annot.data[kegg.annot.data$`Pathway Source`=="KEGG",]
+kegg.gene.sets <- lapply(kegg.annot.data$Genes, 
+                         function(x) unique(unlist(str_split(x, " \\| "))))
+names(kegg.gene.sets) <- kegg.annot.data$`Pathway Id`
+kegg.gene.sets <- kegg.gene.sets[unlist(lapply(kegg.gene.sets, length))>=MIN_SET_SIZE]
+```
+
+Enrichment analysis
+```{r}
+total.genes <- sum(grepl("^PF3D7", rownames(fc2$counts))) #Size of universe
+top_genes_set <- top_ring_ruv
+go.test <- as.data.frame(do.call("rbind",lapply(go.gene.sets, function(gset){
+  set.size <- length(gset)
+  gin_all <- gset[gset %in% top_genes_set$GeneID]
+  gin_up <- gset[gset %in% top_genes_set$GeneID[top_genes_set$logFC>0]]
+  gin_down <- gset[gset %in% top_genes_set$GeneID[top_genes_set$logFC<0]]
+  result <- c(set.size=set.size,
+              n.both=length(gin_all),
+              n.up=length(gin_up),
+              n.down=length(gin_down),
+              p.value.both=phyper(length(gin_all)-1, m=length(gset), n=total.genes-length(gset),
+                                k=nrow(top_genes_set), lower.tail = FALSE),
+              p.value.up=phyper(length(gin_up)-1, m=length(gset), n=total.genes-length(gset),
+                                k=sum(top_genes_set$logFC>0), lower.tail = FALSE),
+              p.value.down=phyper(length(gin_down)-1, m=length(gset), n=total.genes-length(gset),
+                                k=sum(top_genes_set$logFC<0), lower.tail = FALSE)
+                 )
+  return(result)
+})))
+go.test <- go.test[order(go.test$p.value.both),]
+
+go.test$p.value.both <- format(go.test$p.value.both, scientific=TRUE, digits=3)
+go.test$p.value.up <- format(go.test$p.value.up, scientific=TRUE, digits=3)
+go.test$p.value.down <- format(go.test$p.value.down, scientific=TRUE, digits=3)
+kable(go.test)
+```
+
+```{r}
+keg.test <- as.data.frame(do.call("rbind",lapply(kegg.gene.sets, function(gset){
+  set.size <- length(gset)
+  gin_all <- gset[gset %in% top_genes_set$GeneID]
+  gin_up <- gset[gset %in% top_genes_set$GeneID[top_genes_set$logFC>0]]
+  gin_down <- gset[gset %in% top_genes_set$GeneID[top_genes_set$logFC<0]]
+  result <- c(set.size=set.size,
+              n.both=length(gin_all),
+              n.up=length(gin_up),
+              n.down=length(gin_down),
+              p.value.both=phyper(length(gin_all)-1, m=length(gset), n=total.genes-length(gset),
+                                k=nrow(top_genes_set), lower.tail = FALSE),
+              p.value.up=phyper(length(gin_up)-1, m=length(gset), n=total.genes-length(gset),
+                                k=sum(top_genes_set$logFC>0), lower.tail = FALSE),
+              p.value.down=phyper(length(gin_down)-1, m=length(gset), n=total.genes-length(gset),
+                                k=sum(top_genes_set$logFC<0), lower.tail = FALSE)
+                 )
+  return(result)
+})))
+keg.test <- keg.test[order(keg.test$p.value.both),]
+keg.test$Pathway <- kegg.annot.data$Pathway[match(rownames(keg.test),
+                                                  kegg.annot.data$`Pathway Id`)]
+kable(keg.test, digits = 4)
+```
+
 ##Gene sets from Daily et al and Shock et al
@@
 
-##KEGG analysis
+###Summary Gene Set Plot
 ```{r}
-# setwd(wd)
-# 
-# entrez_de <- entrez_geneID_mappings$`[Entrez Gene ID]`[entrez_geneID_mappings$`[Gene ID]` %in% top_ring_ruv$GeneID]
-# entrez_de <- entrez_de[entrez_de!="null"]
-# entrez_de <- str_split_fixed(entrez_de, pattern = ",", n = 2)[,1]
-# 
-# kegg <- kegga(entrez_de, species.KEGG="pfa", convert=TRUE)
-# kable(topKEGG(kegg, n=Inf))
-```
-
-##GO analysis
-```{r}
-# go <- goana(entrez_de, species.KEGG="pfa", convert=TRUE)
-# kable(topGO(go, n=100))
+summary.set.data <- fread("./data/genes_in_set_for_figure2.csv",
+                          data.table = FALSE)
+gg <- ggplot(summary.set.data, aes(x=Gene, y=`adj-p`, fill=logFC))
+# gg <- gg + geom_col() + facet_wrap(~SetName, ncol=1, scales = "free_y", 
+#                                                   strip.position = "left")
+gg <- gg + geom_col() + facet_grid(SetName ~ ., space = "free_y", scales = "free_y", switch = "y")
+gg <- gg + coord_flip() 
+gg <- gg + scale_fill_gradient2(low="#0571b0", high = "#ca0020", mid="#ffffff")
+gg <- gg + scale_y_continuous(limits = c(0, 0.15))
+gg <- gg + theme_bw(base_size = 8) + theme(strip.text.y=element_text(angle=180))
+gg <- gg + theme(axis.text.y=element_text(size=4))
+gg
 ```

### 2018-01-28 domain_level/scripts/get_domain_annotations.py

@@ -1,0 +1,128 @@
+import sys
+from collections import defaultdict
+import numpy as np
+from mungo.fasta import FastaReader
+from mungo.sequence import sixFrameTranslation
+
+
+hmmfile = sys.argv[1]
+prefix = sys.argv[2]
+DNA_file = sys.argv[3]
+PROT_file = sys.argv[4]
+
+#Load DNA fasta
+dna_seqs = {}
+for h,s in FastaReader(DNA_file):
+  h = h.split()[0]
+  dna_seqs[h] = s
+
+#Load Protein File
+prot_seqs = {}
+for h,s in FastaReader(PROT_file):
+  if len(h.split()) > 2:
+    h = h.split()[0] + "_" + h.split()[-1]
+  else:
+    h = "_".join(h.split())
+  prot_seqs[h] = s
+
+def overlap(start1, end1, start2, end2):
+    """Does the range (start1, end1) overlap with (start2, end2)?"""
+    return not (end1 < start2 or end2 < start1)
+
+transcripts = defaultdict(list)
+with open(hmmfile, 'r') as infile:
+  for line in infile:
+    if line[0]=="#":
+      continue
+    line=line.split()
+    tran = line[0]
+    if line[-1]!="-":
+      tran = tran + "_" + line[-1]
+    dom = line[3]
+    E = float(line[12])
+    from_ = int(line[17])
+    to = int(line[18])
+
+    transcripts[tran].append((dom, from_, to, E))
+
+annotations = {}
+for tran in transcripts:
+  #sort by evalue
+  keep = []
+  transcripts[tran] = sorted(transcripts[tran], key=lambda tup: tup[3])
+  for dom in transcripts[tran]:
+    isOverlap=False
+    for k in keep:
+      if overlap(dom[1],dom[2],k[1],k[2]):
+        isOverlap=True
+        break
+    if not isOverlap:
+      keep.append(dom)
+  annotations[tran] = keep
+
+with open(prefix+"_hits.txt",'w') as outfile:
+  for tran in annotations:
+    annotations[tran] = sorted(annotations[tran], key=lambda tup: tup[1])
+    #outfile.write(",".join([tran]+[tup[0].split("_")[-1] for tup in annotations[tran]])+"\n")
+    outfile.write(tran+"\t"+",".join([str(tup) for tup in annotations[tran]])+"\n")
+
+
+
+def findPosition(protein, dna):
+  translated = sixFrameTranslation(dna)
[…]

### 2018-01-28 domain_level/scripts/hierarchical_usearch.py

@@ -1,0 +1,133 @@
+from mungo.fasta import FastaReader
+from mungo.sequence import sixFrameTranslation
+import sys, os
+import shutil
+from optparse import OptionParser
+from subprocess import check_call
+from collections import defaultdict
+import subprocess
+
+
+USEARCH = "/home/users/allstaff/tonkin-hill.g/var_blocks/third-party/usearch8"
+
+
+def run_usearch(inputfile, outdir, minI, cpu, verbose):
+
+  prefix = os.path.splitext(os.path.basename(inputfile))[0]
+
+  #copy file to output directory and rename transcripts with temprary
+  # place folders
+  names = {}
+  segCount = 0
+  with open(outdir + prefix + "_renamed.fa", 'w') as outfile:
+    for h,s in FastaReader(inputfile):
+      names["block_" + str(segCount)] = "_".join(h.split())
+      outfile.write(">" + "block_" + str(segCount) + "\n")
+      s = s.replace("*", "X")
+      outfile.write(s + "\n")
+      segCount += 1
+
+  inputfile = outdir + prefix + "_renamed.fa"
+
+  #read in sequences
+  sequences = []
+  for h,s in FastaReader(inputfile):
+    sequences.append((h,s))
+
+  num_sequences = len(sequences)
+  sub_size = num_sequences/cpu + 1
+
+  #write out into seperate files
+  files = []
+  for i in range(0,cpu):
+    filename = outdir + prefix + "_query" + str(i) + ".fa"
+    with open(filename, 'w') as outfile:
+      for seq in sequences[i*sub_size:(i+1)*sub_size]:
+        outfile.write(">"+seq[0]+"\n")
+        outfile.write(seq[1]+"\n")
+    files.append(filename)
+
+  #now run usearch with each query file
+  processes = set()
+  #-usearch_global ./test.fa -db ./param_p1e-5_k5_radiusThresh1.fa -id 0.2 -maxaccepts 0 -maxrejects 0
+  #-fastapairs alignFasta.out -uc uc.out  -qmask none -dbmask none -wordlength 2 -alpha aa -fulldp -threads 20
+  for f in files:
+    out = f.strip(".fa") + "_usearch.tab"
+    search_cmd = (USEARCH
+      + " -usearch_global " + f
+      + " -db " + inputfile
+      + " -id " + str(minI)
+      + " -maxaccepts 0 -maxrejects 0"
+      + " -qmask none -dbmask none"
[…]

### 2018-01-28 domain_level/scripts/translate_transcripts_with_HMM.py

@@ -1,0 +1,229 @@
+from mungo.sequence import sixFrameTranslation
+from mungo.fasta import FastaReader
+import os, sys
+from optparse import OptionParser
+from subprocess import check_call
+from collections import defaultdict
+
+HMMER3 = "/usr/local/bioinfsoftware/hmmer/hmmer-3.1b1/bin/hmmsearch"
+
+def split_easy_from_hard(inputfile, outputdir, num_stops):
+  seqCount = 0
+  badSeqs = 0
+  bad_lengths = []
+
+  output_file = (outputdir + os.path.splitext(os.path.basename(inputfile))[0]
+    + "_translatedL2stops.fa")
+
+  with open(output_file + "_BadSeqs", 'w') as badfile:
+    with open(output_file, 'w') as outfile:
+
+      for h,s in FastaReader(inputfile):
+        stops = 9999
+        translation = sixFrameTranslation(s)
+
+        for frame in translation:
+          st = translation[frame].count('*')
+          if st < stops:
+            best = frame
+            stops = st
+
+        if stops <= num_stops:
+          outfile.write(">" + h + " frame_" + str(best) + "\n")
+          outfile.write(translation[best] + "\n")
+        else:
+          badSeqs += 1
+          bad_lengths.append(len(s))
+          badfile.write(">" + h + "\n")
+          badfile.write(s + "\n")
+
+        seqCount += 1
+  print (str((100.0*badSeqs)/seqCount) + "percent or "
+    + str(badSeqs) + " out of " + str(seqCount) + " were not translated.")
+
+  return output_file, output_file + "_BadSeqs"
+
+def get_long_ORFS(translation, len_cutoff):
+  keep = []
+  for frame in translation:
+    orfs = translation[frame].split('*')
+    o_counts = 0
+    for o in orfs:
+      if len(o) >= len_cutoff:
+        keep.append(("_frm"+str(frame)+"_orf"+str(o_counts), o))
+      o_counts += 1
+
+  return keep
+
+def pull_out_long_ORFs(bad_file, len_cutoff, outputdir):
+
+  out_file = (outputdir + os.path.splitext(os.path.basename(bad_file))[0]
+    + "_TranslongORFS.fa")
+
+  num_seqs = 0
+  num_orfs = 0
+  with open(out_file, 'w') as outfile:
+    for h,s in FastaReader(bad_file):
+      num_seqs += 1
+      translation = sixFrameTranslation(s)
+      orfs = get_long_ORFS(translation, len_cutoff)
+      for o in orfs:
[…]

### 2018-01-28 domain_level/scripts/get_domain_annotations.py
small change to script

@@ -104,12 +104,15 @@ with open(prefix+"_domains.fasta", 'w') as outfas:
     for tran in annotations:
       dna_h = tran.split("_f")[0]
       dna = dna_seqs[dna_h]
+      dom_num=0
       for tup in annotations[tran]:
-        outfas.write(">" + tran + "_" + tup[0] + "\n")
+        dom_num += 1
+        domname = tran + "_" + tup[0] + "_domNum" + str(dom_num)
+        outfas.write(">" + domname + "\n")
         prot = prot_seqs[tran][tup[1]-1:tup[2]]
         outfas.write(prot + "\n")
         dna_pos = findPosition(prot, dna)
-        outSAF.write("\t".join([tran + "_" + tup[0], dna_h, str(dna_pos[0]), str(dna_pos[1]), "-"])+"\n")
+        outSAF.write("\t".join([domname, dna_h, str(dna_pos[0]), str(dna_pos[1]), "-"])+"\n")
 
 
 
